\chapter{Ulrich's MA 553 Problems for Spring `16}
\section{Homework 1}
\begin{problem}
Let $G$ be a group, $a\in G$ an element of finite order $m$, and $n$ a
positive integer. Prove that
\[
|a^n|=\frac{m}{\gcd(m,n)}.
\]
\end{problem}
\begin{proof}

\end{proof}

\begin{problem}
Let $G$ be a group, and let $a$, $b$ be elements of finite order $m$, $n$
respectively. Show that if $ba=ab$ and $\langle a\rangle\cap\langle
b\rangle=\{e\}$, then $|ab|=\lcm(m,n)$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G$ be a group and $H$, $K$ normal subgroups with $H\cap K=\{e\}$. Show
that
\begin{enumerate}[label=(\alph*)]
\item $hk=kh$ for every $h\in H$, $k\in K$.
\item $HK$ is a subgroup of $G$ with $HK\cong H\times K$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Show that $A_4$ has no subgroup of order $6$ (although $6\mid 12=|A_4|$).
\end{problem}
\begin{proof}
\end{proof}

\section{Homework 2}
\begin{problem}
Let $G$ be the group of order $2^3\cdot 3$, $n\geq 2$. Show that $G$ has a
normal $2$-subgroup $\neq\left\{e\right\}$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G$ be a group of order $p^2q$, $p$ and $q$ primes. Show that the Sylow
$p$-Sylow subgroup or the $q$-Sylow subgroup of $G$ is normal in $G$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G$ be a subgroup of order $pqr$, $p<q<r$ primes. Show that the
$r$-Sylow subgroup of $G$ is normal in $G$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G$ be a group of order $n$ and let $\varphi\colon G\to S_n$ be given by
the action of $G$ on $G$ via translation.
\begin{enumerate}[label=(\alph*)]
\item For $a\in G$ determine the number and the lengths of the disjoint
  cycles of the permutation $\phi(a)$.
\item Show that $\varphi(G)\nsubset A_n$ if and only if $n$ is even and $G$
  has a cyclic $2$-Sylow subgroup.
\item If $n=2m$, $m$ odd, show that $G$ has a subgroup of index $2$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Show that the only simple groups $\neq\left\{e\right\}$ of order $<60$ are
the groups of prime order.
\end{problem}
\begin{proof}
\end{proof}

\section{Homework 3}
\begin{problem}
Let $G$ be a finite group, $p$ a prime number, $N$ the intersection of all
$p$-Sylow subgroups of $G$. Show that $N$ is a normal $p$-subgroup of $G$
and that every normal $p$-subgroup of $G$ is contained in $N$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G$ be a group of order $231$ and let $H$ be an $11$-Sylow subgroup of
$G$. Show that $H\subset Z(G)$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $G=\left\{e,a_1,a_2,a_3\right\}$ be a non-cyclic group of order $4$ and
define $\varphi\colon S_3\to\Aut(G)$ by $\varphi(\sigma)(e)=e$ and
$\varphi(\sigma)(a_1)=a_{\sigma(i)}$. Show that $\varphi$ is well-defined and an
isomorphism of groups.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Determine all groups of order $18$.
\end{problem}
\begin{proof}
\end{proof}

\section{Homework 4}
\begin{problem}

\end{problem}

\section{Homework 5}


\section{Homework 6}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA553-Quals"
%%% End:
