\begin{problem}
For $I$ an $R$-ideal consider the multiplicatively
closed set $S=1+I$. Show that
\begin{enumerate}[noitemsep,label=(\alph*)]
\item $\tilde S=R\setminus\bigcup\mathfrak{m}$, where the
  union is taken over all
  $\mathfrak{m}\in\MSpec(R)\cap V(I)$.
\item $\MSpec(S^{-1}R)$ and $\MSpec(R/I)$ are homeomorphic.
\end{enumerate}
\end{problem}
\begin{proof}
(a) By 4.19, we have
\[
\tilde
S=R\setminus\bigcup_{\substack{\mathfrak{p}\in\Spec(R)\\\mathfrak{p}\cap
  S=\emptyset}}\mathfrak{p}.
\]
But $\mathfrak{p}\cap S=\mathfrak{p}\cap (1+I)=\emptyset$ if and
only if $\mathfrak{p}+I\neq R$ if and only if there is some
maximal ideal $\mathfrak{m}\supset \mathfrak{p}+I$.

For the former equivalence: $\implies$ Suppose that
$\mathfrak{p}\cap S=\mathfrak{p}\cap(1+I)=\emptyset$, then if
$\mathfrak{p}+I=R$ for some $x\in\mathfrak{p}$, $y\in I$ we have
$x+y=1$. But then $x=1-y\in\mathfrak{p}\cap S$; this is a
contradiction. $\impliedby$ Conversely, if
$\mathfrak{p}\cap S\neq\emptyset$, $x=1+y\in\mathfrak{p}$ for some
$y\in I$ so $x-y=(1+y)-y=1\in\mathfrak{p}+I$ implies
$\mathfrak{p}+I=R$.

For the latter equivalence: $\implies$ Suppose $\mathfrak{p}+I\neq R$,
then $\mathfrak{p}+I$ is a proper ideal of $R$ so, by 1.5, is
contained in a maximal ideal $\mathfrak{m}$. $\impliedby$ Conversely,
if $\mathfrak{m}\subsetneq R$ is a maximal ideal containing
$\mathfrak{p}+I$ then $\mathfrak{p}+I\neq R$ for otherwise
$\mathfrak{m}=R$. Then it suffices to take the union over all maximal
ideals $\mathfrak{m}\supset I$.
\\\\
(b) We will show that $\MSpec(S^{-1}R)\approx\MSpec(R)\cap V(I)$ and
$\MSpec(R/I)\approx\MSpec(R)\cap V(I)$ so that, by the transitivity of
homeomorphism, we have $\MSpec(S^{-1}R)\approx\MSpec(R/I)$. By
4.21(a), $\Spec(R/I)\approx V(I)$ so the restriction
$\MSpec(R/I)\approx\MSpec(R/I)\cap V(I)$. To see that
$\MSpec(S^{-1}R)\approx\MSpec(R)\cap V(I)$, let $\varphi\colon R\to
S^{-1}R$ be the canonical homomorphism sending $x\mapsto x/1$, then
$\phi$ induces a continuous closed map
$^a\phi\colon\Spec(S^{-1}R)\to\Spec(R)$ taking
$\bar{\mathfrak{p}}\mapsto\mathfrak{p}$, i.e., ideal extension. Thus,
by 4.13(d), there is a one-to-one correspondence between
$\bar{\mathfrak{p}}\in\Spec(S^{-1}M)$ and its extension
$\mathfrak{p}\in\Spec(R)$ with $\mathfrak{p}\cap S=\emptyset$ so that
it suffices to show that $^a\phi(\MSpec(S^{-1}R))=\MSpec(R)\cap
V(I)$. But this is easy: If $\bar{\mathfrak{m}}\in\MSpec(S^{-1}R)$
then its contraction is a maximal ideal $\mathfrak{m}\supset I$ by
part (a), hence is in $\MSpec(R)\cap V(I)$. Conversely, if
$\mathfrak{m}\in\MSpec(R)\cap V(I)$, again, by part (a),
$\mathfrak{m}$ is a maximal ideal not meeting $S$ so that by 4.13(d),
there exist some maximal ideal $\bar{\mathfrak{m}}$ contracting to
$\mathfrak{m}$. It follows that $\MSpec(S^{-1}R)\approx\MSpec(R/I)$.
\end{proof}
\newpage
\begin{problem}
Show that the following are equivalent for a ring $R$:
\begin{enumerate}[noitemsep,label=(\alph*)]
\item there exist rings $R_1\neq 0$ and $R_2\neq 0$ so that
  $R\cong R_1\times R_2$;
\item there exist an idempotent $e\in R$ with $e\neq 0$ and
  $e\neq 1$;
\item $\Spec(R)$ is disconnected.
\end{enumerate}
\end{problem}
\begin{proof}
(a) $\iff$ (b) is immediate for suppose $R\cong R_1\times
R_2$ by $\phi\colon R\to R_1\times R_2$. Then, since $\phi$ is a
bijection, there exist an $r\in R$ that maps to the idempotent
element $(1,0)\in R_1\times R_2$.

Conversely, suppose $e\in R$ is idempotent. Then $e'=1-e$ is also
idempotent since
\[
(e')^2=(1-e)^2=1-2e+e^2=1-2e+e=1-e.
\]
Moreover
\[
ee'=e(1-e)=e-e^2=e-e=0.
\]
Let $R_1$ and $R_2$ be the subrings of $R$ generated by $e$ and
$e'$, respectively. Then we claim that $R\cong R_1\times R_2$ via
$\phi(r)=(re,re')$. It is clear that $\phi$ is a ring
homomorphism: take $r_1,r_2\in R$ then
\begin{align*}
\phi(r_1+r_2)&=((r_1+r_2)e,(r_1+r_2)e')&
\phi(r_1r_2)&=(r_1r_2e,r_1r_2e')\\
&=(r_1e+r_2e,r_1e'+r_2e')&
&=\left(r_1r_2e^2,r_1r_2(e')^2\right)\\
&=(r_1e,r_1e')+(r_2e,r_2e')&
&=(r_1e,r_1e')(r_2e,r_2e')\\
&=\phi(r_1)+\phi(r_2)&
&=\phi(r_1)\phi(r_2).
\end{align*}
To prove surjective take $(r,s)\in R_1\times
R_2$ then, $r=r_1e$ and $s=r_2e'$ for $r_1,r_2\in R$ then
\begin{align*}
\phi(r_1e+r_2e')&=\phi(r_1e)+\phi(r_2e')\\
&=(r_1e,r_1ee')+(r_2e'e,r_2ee')\\
&=(r_1e,0)+(0,r_2e')\\
&=(r_1e,r_2e')\\
&=(r,s).
\end{align*}
To prove injectivity take $r\in\ker\phi$. Then
$\phi(r)=(re,re')=(0,0)$. Then $re-re'=r(e-e')=r\cdot 1=0$ so
$r=0$.

(a) $\implies$ (c) Recall that a topological space $X$ is
disconnected if there exist disjoint open sets $A,B$ with
$X=A\cup B$. Suppose $R\cong R_1\times R_2$. Then
$\Spec(R)\approx \Spec(R_1\times R_2)$: Keeping the notation as
before, $\phi$ is a set bijection so it induces a bijection, call
it $\phi^*$, on $\Spec(R)\to\Spec(R_1\times R_2)$ by sending
$\Spec(I)\mapsto\Spec(\phi(I))$; Now let $I\subset R$ be an
ideal, then
\[
\phi^*(V(I))=\phi^*\left(V(eI+e'I)\right)=V(\phi(eI)+\phi(e'I))=V(eI\times e'I)
\]
is closed. Thus, $\phi^*$ is a homeomorphism. Now, we claim that
the sets $A=V(R_1\times 0)$ and $B=V(0\times R_2)$ constitute a
separation of $R$. First note by 4.20(2) that
\[
A\cup B=V(R_1\times 0)\cup V(0\times R_2)=V((R_1\times
0)\cap(0\times R_2))=V(0)=\Spec(R).
\]
Moreover
\[
A\cap B=V(R_1\times 0)\cap V(0\times R_2)=V(R_1\times 0+0\times R_2)=V(R)=\emptyset.
\]
Since both $A$ and $B$ are closed, it follows that $A,B$ forms a
separation of $\Spec(R)$.
\\\\
(c) $\implies$ (b)
As suggested by Matsumura: If $\Spec(R)=V(I_1)\cup V(I_2)$ with
$V(I_1)\cap V(I_2)\cap=\emptyset$, then $I_1+I_2=R$ and
$I_1I_2\subset\nil(R)$. So $1=e_1+e_2$ and $(e_1e_2)^n=0$ for
some $e_1\in I_1$, $e_2\in I_2$. So we have that
\[
1=(e_1+e_2)^{2n}=e_1^nx_1+e_2^nx_2
\]
with $x_1,x_2\in R$. So $e=e_1^nx_1$ satisfies $e(1-e)=0$, i.e.,
$e$ is idempotent.
\end{proof}
\newpage
\begin{problem}
A topological space is called \emph{Noetherian} if the set of
closed sets satisfies the dcc. Show that if a ring $R$ is
Noetherian then so is $\Spec(R)$, but that the converse does not
hold.
\end{problem}
\begin{proof}
We will first prove the following useful results:
\begin{lemma*}
Let $R$ be a commutative ring with identity. Then
\begin{enumerate}[noitemsep,label=(\roman*)]
\item $V(I)=V(\sqrt{I})$.
\item $I\subset J$ implies $V(I)\supset V(J)$.
\item $V(I)\supset V(J)$ implies $\sqrt{I}\subset\sqrt{J}$.
\end{enumerate}
\end{lemma*}
\begin{proof}[Proof of lemma]
\renewcommand\qedsymbol{$\clubsuit$}
(i) It is clear that for every prime ideal
$\mathfrak{p}\supset\sqrt{I}$ we have $\mathfrak{p}\supset I$ so
it suffice to prove that if $\mathfrak{p}\supset I$ then
$\mathfrak{p}\supset\sqrt{I}$. But this is clear since if
$x\in\sqrt{I}$ then $x^k\in I$ for some positive integer $k$ so
$x^k\in\mathfrak{p}$ and since $\mathfrak{p}$ is prime
$x\in\mathfrak{p}$. Thus, $V(I)=V(\sqrt{I})$.
\\\\
(ii) Suppose $I\subset J$. Then every prime ideal
$\mathfrak{p}\supset J$ must also contain $I$. Thus, $V(I)\supset
V(J)$.
\\\\
(iii) Suppose $V(I)\supset V(J)$. Then, for every prime ideal
$\mathfrak{p}\supset J$, $\mathfrak{p}\supset I$ so
\[
\sqrt{J}=\bigcap_{\mathfrak{p}\supset J}\mathfrak{p}\supset
\bigcap_{\mathfrak{p}\supset
    J}\mathfrak{p}\cap\bigcap_{\substack{\mathfrak{q}\supset
    I\\\mathfrak{q}\nsupset J}}\mathfrak{q}=\sqrt{I}.\qedhere
\]
\end{proof}
It suffices to reduce to the case of varieties of ideals in
$R$ since varieties generate the Zariski topology on
$\Spec(R)$. Suppose
\[
V(I_1)\supset V(I_2)\supset\cdots
\]
is a descending chain of varieties in $\Spec(R)$. Then, by the (iii)
of the lemma and the nullstellensatz, the latter chain is in
one-to-one correspondence with the ascending chain of radical ideals
\[
\sqrt{I_1}\subset\sqrt{I_2}\subset\cdots
\]
which must stabilize since $R$ is Noetherian. It follows that the
chain $V(I_1)\supset V(I_2)\supset\cdots$ stabilizes so $\Spec(R)$ is
Noetherian.
\end{proof}
\newpage
\begin{problem}
A nonempty closed subset $V$ of a topological space is called
\emph{irreducible} if $V=V_1\cup V_2$, $V_1$ and $V_2$ closed
subset, implies $V_1=V$ or $V_2=V$.
\begin{enumerate}[noitemsep,label=(\alph*)]
\item Show that in a Noetherian topological space every nonempty
  closed subset is a finite union of irreducible closed subsets.
\item Show that $V(\mathfrak{p})$, $\mathfrak{p}\in\Spec(R)$, are
  exactly the irreducible closed subsets of $\Spec(R)$.
\end{enumerate}
\end{problem}
\begin{proof}
(a) Let $X$ be a Noetherian topological space. Let
\[
\Lambda=\left\{ \,V\subset X\;\middle|\;\text{$V$ is closed and not a
    finite union of irreducible closed subsets} \right\}.
\]
Then, by the dcc, $\Lambda$ contains a minimal element, say $W$. Then
$W$ is not irreducible so we can write $W=W_1\cup W_2$ where $W_1\neq
W$ and $W_2\neq W$. By minimality of $W$, $W_1$ and $W_2$ are finite
unions of irreducible closed subsets so $W_1=\bigcup_{i=1}^k
W_1^{(i)}$ and $W_2=\bigcup_{i=1}^\ell W_2^{(i)}$ so
\[
W=W_1\cup W_2=\left( \bigcup_{i=1}^\ell W_1^{(i)} \right)\cup\left(
  \bigcup_{i=1}^k W_2^{(i)} \right)
\]
a contradiction. Thus, every closed subset $V$ can be expressed as the
finite union of irreducible closed subsets.
\\\\
(b) We prove the contrapositive. Suppose that $I\subset R$ is not
prime. Then we can find $x,y\in R$ with $xy\in I$, but $x\notin I$,
$y\notin I$. Thus,
\[
V((I,x))\cup V((I,y))=V((I,x)\cap (I,y))=V(I),
\]
but neither $V((I,x))\neq V(I)$ or $V((I,y))\neq V(I)$ so $V(I)$ is
not irreducible. \qedhere
% (b) Suppose that $V(\mathfrak{p})=V_1\cup V_2$. Let
% $I=\bigcap_{\mathfrak{p}\in V_1}$ and $J=\bigcap_{\mathfrak{p}\in
%   V_1}\mathfrak{p}$. Then $V(I)=V_1$ and $V(J)=V_2$ by the
% nullstellensatz. Thus,
% \[
% V_1\cup V_2=V(I)\cup V(J)=V(IJ)=V(\mathfrak{p})
% \]
% so $IJ=\mathfrak{p}$ implies $I\subset\mathfrak{p}$ or
% $J\subset\mathfrak{p}$. But $V(I)\subset V(\mathfrak{p})$,
% $V(J)\subsetV(\mathfrak{p})$ implies $I\supset\mathfrak{p}$ and
% $J\supset\mathfrak{p}$ so $I=\mathfrak{p}$ or $J=\mathfrak{p}$. Hence
% $V(\mathfrak{p})=V_1$ or $V(\mathfrak{p})=V_2$ and we have that
% $V(\mathfrak{p})$ is irreducible.
\end{proof}
\newpage
\begin{problem}
Show that a Noetherian ring has only finitely many minimal prime
ideals.
\end{problem}
\begin{proof}
Since $R$ is Noetherian, by Problem 5.3, $\Spec(R)$ is Notherian, that
is, it satisfies the dcc. Thus, by Problem 5.4, $\Spec(R)$ is the
union of finitely many irreducible subsets
$V(0)=\bigcup_{i=1}^nV(\mathfrak{p}_i)$ where $\mathfrak{p}_i$ is
prime. Now, suppose $\mathfrak{q}$ is a minimal prime (we are
guaranteed one if $R\neq 0$ by the next problem). Then
$\mathfrak{q}\in V(0)$ so $\mathfrak{q}\in V(\mathfrak{p}_i)$ for some
$1\leq i\leq n$. Then $\mathfrak{q}\supset\mathfrak{p}_i$, but by
minimality $\mathfrak{q}=\mathfrak{p}_i$. It follows that $R$ contains
$\leq n$ minimal prime ideals, in particular, finitely many.
\end{proof}
\newpage
\begin{problem}
Show that a nonzero ring has at least one minimal prime ideal.
\end{problem}
\begin{proof}
% \begin{lemma*}
% Let $\left\{\mathfrak{p}_\alpha\right\}\subset R$ be a chain of prime
% ideals ordered by reverse-inclusion. Then $\bigcap\mathfrak{p}_\alpha$ is prime.
% \end{lemma*}
% \begin{proof}[Proof of lemma]
% \renewcommand\qedsymbol{$\clubsuit$}
% Put $\mathfrak{p}=\bigcap\mathfrak{p}_\alpha$. Suppose
% $xy\in\mathfrak{p}$. Then $xy\in\mathfrak{p}_\alpha$ for every
% $\alpha$. Fix an $\alpha$ and, without loss of generality, suppose
% $x\in\mathfrak{p}_\alpha$. Let $\mathfrak{p}_\beta$ be another prime
% ideal. Then $\mathfrak{p}_\beta\subset\mathfrak{p}_\alpha$ or
% $\mathfrak{p}_\beta\supset\mathfrak{p}_\alpha$. If
% $\mathfrak{p}_\beta\supset\mathfrak{p}_\alpha$, we are done
% ($x\in\mathfrak{p}_\beta$). Otherwise, suppose
% $x\notin\mathfrak{p}_\beta$. Then $y\in\mathfrak{p}_\beta$ so
% $y\in\mathfrak{p}_\beta$. In either case, $\mathfrak{p}$ is
% prime.
% \end{proof}
% Since R is nonzero, it contains a maximal ideal which is a
% prime ideal. Thus the set A of all prime ideals of R is
% nonempty. A is ordered by reverse-inclusion. Let A be a totally
% ordered subset of A. It's pretty clear that J=⋂I∈AI is in fact
% an ideal. Not so clear, however, is that it is prime. Let
% xy∈J. Then xy∈I for all I∈A. Now let B={I∈A|y∈I}. Let
% K=⋂I∈BI. Since A is totally ordered, either K=J (and we're
% done, since then y∈J) or K⊃J and for all I∈A such that I is
% properly contained in K, we have y∉I. But that means that for
% all those I,x∈I, since they are prime. Hence x∈J. In either
% case, J is prime as desired. Hence by Zorn's lemma we get a
% maximal element which in this case is a minimal prime ideal.
We will proceed by Zorn's lemma. Let $\Lambda$ be the set of all
prime ideals of $R$. This set is nonempty by 1.4 since every
nonzero ring contains a maximal ideal, in particular, a prime
ideal. $\Lambda$ is ordered by reverse-inclusion. Let
$L\subset\Lambda$ be a totally ordered subset of $\Lambda$. We
claim that $\mathfrak{q}=\bigcap_{\mathfrak{p}\in L}\mathfrak{p}$
is a prime ideal: It is clear that $\mathfrak{q}$ is an
ideal. Suppose $xy\in\mathfrak{q}$. Then
$\mathfrak{xy}\in\mathfrak{p}$ for every $\mathfrak{p}\in L$. Let
$K$ be the collection of all $\mathfrak{p}\ni x$. Then either
$\bigcap_{\mathfrak{p}\in K}\mathfrak{p}=\mathfrak{q}$ (in which
case we are done) or $\bigcap_{\mathfrak{p}\in
  K}\mathfrak{p}\supset\mathfrak{q}$ where
$x\notin\mathfrak{q}$. But then, since every $\mathfrak{p}\in L$
is prime, they must contain $x$ so $x\in\bigcap_{\mathfrak{p}\in
  K}\mathfrak{p}$. Thus, $\mathfrak{q}$ is prime. By Zorn's
lemma, it follows that $\Lambda$ contains a minimal element, in
this case, a smallest (by reverse-inclusion) prime ideal.
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA557-HW-Current"
%%% End:
