\begin{problem}
Let $R$ be a ring such that $R_{\mathfrak{m}}$ is Noetherian for
every $\mathfrak{m}\in\MSpec R$ and for every $f\in R$, $f\neq
0$, there exists at most finitely many $\mathfrak{m}\in\MSpec R$
with $f\in\mathfrak{m}$. Show that $R$ is Noetherian.
\end{problem}
\begin{proof}
As suggested by Atiyah and MacDonald, we will attempt to show that if
$I$ is a proper ideal of $R$ it is finitely generated. Since $I\subsetneq
R$ is proper, there are finitely many maximal ideals which contain it since
given any point $f\in I$ there are only finitely many ideals which
contain $f$. Let $\mathfrak{m}_1,...,\mathfrak{m}_r$ be the maximal
ideals containing $I$ and let $\mathfrak{m}_1,...,\mathfrak{m}_{r+s}$ be
the finitely many maximal ideals which contain $f\in I$. Supposing
$\mathfrak{m}_{r+1},...,\mathfrak{m}_{r+s}$ do not contain $I$, then take
$x_j\in I\smallsetminus\mathfrak{m}_j$ for $r<j\leq r+s$. Since each
$R_{\mathfrak{m}_i}$ is Noetherian, the extension $I_{\mathfrak{m}_i}$ is
finitely generated in $R_{\mathfrak{m}_i}$, say by $x_i^j$ for $1\leq i\leq
r$ and $1\leq j\leq n_i$ where $n_i$ depends on $i$. Then the ideal
$J\coloneqq (x_0,\left\{x_i^j\right\},x_{r+1},...,x_{r+s})$ and $I$
coincide on $R_{\mathfrak{m}}$ for all $\mathfrak{m}\in\MSpec R$. Therefore
$J=I$.
\end{proof}
\newpage
\begin{problem}
Let $k$ be a field, $T=k\left[\bigl\{X_j\;\middle|\;j\in\NN\bigr\}\right]$ a
polynomial ring, $n_i$ a sequence in $\NN$ with
$0<n_i-n_{i-1}<n_{i+1}-n_i$ for every $i$,
$\mathfrak{p}_i=\left(X_j\;\middle|\;n_i\leq j\leq
  n_{i+1}-1\right)$,
$W=T\smallsetminus\bigcup_{i\in\NN}\mathfrak{p}_i$ and
$R=W^{-1}T$. Show that
\begin{enumerate}[label=(\alph*)]
\item $\MSpec
  R=\left\{\,W^{-1}\mathfrak{p}_i\;\middle|\;i\in\NN\,\right\}$
  and $R_{W^{-1}\mathfrak{p}_i}$ is a Noetherian ring of
  dimension $n_{i+1}-n_i$.
\item $R$ is a Noetherian ring and $\dim R=\infty$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $k$ be a field, $R=k[X_1,...,X_n]$ a polynomial ring and
$f\in R\smallsetminus k$. Show that $\dim R/(f)=n-1$.
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $R$ be a Noetherian semilocal ring and $M$ a finite
$R$-module.
\begin{enumerate}[label=(\alph*)]
\item Write
  $n=\max\left\{\,\mu_{R_{\mathfrak{m}}}(M_{\mathfrak{m}})\;\middle|\;\mathfrak{m}\in\MSpec
  R\,\right\}$. Show that $M$ can be generated by $n$ elements.
\item Assume $R$ is a domain. Show that if $M$ is projective then
  $M$ is free.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $R$ be a Noetherian ring. Show that the following are
equivalent:
\begin{enumerate}[label=(\roman*)]
\item $R$ is a normal ring
\item $R$ is a finite direct product of normal domains
\item $R$ is a reduced and integrally closed in its total ring of
  quotients.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA557-HW-Current"
%%% End:
