\begin{problem}
Let $R$ be a Noetherian ring and $I,J$ $R$-ideals. Write
$I^{\langle  J \rangle}=\bigcup_{n\geq 1}(I:J^n)$, which is
called the \emph{saturation of $I$ with respect to $J$}. Show:
\begin{enumerate}[label=(\alph*)]
\item If $I=\bigcap_{i=1}^m\mathfrak{q}_i$ with $\mathfrak{q}_i$
  $\mathfrak{p}_i$-primary, then $I^{\langle
    J\rangle}=\bigcap_{J\nsubset\mathfrak{p}_i}\mathfrak{q}_i$.
\item $I^{\langle  J \rangle}$ is the unique largest $R$-ideal
  that coincides with $I$ locally on the open set
  $\Spec(R)\smallsetminus V(J)$.
\end{enumerate}
\end{problem}
\begin{proof}
(a) We shall demonstrate double inclusion: Let
$\bigcap_{i=1}^m\mathfrak{q}_i$ be a minimal decomposition of $I$ into
primary ideals where $\mathfrak{q}_i$ is
$\mathfrak{p}_i$-primary. $\implies$ Suppose $x\in I^{\langle J\rangle}$
then $xJ^n\subset I$ for some $n\geq 1$. Given $i$ such that
$\mathfrak{p}_i\nsupset J$\footnote{\textcolor{Red}{Why does such
    an ideal exist? Well, suppose that $\mathfrak{p}_i\supset J$
    for all $1\leq i\leq   m$. Then
  $J\subset\bigcap_{i=1}^m\mathfrak{p}_i=\bigcap_{i=1}^m\sqrt{\mathfrak{q}_i}=\sqrt{\bigcap_{i=1}^m\mathfrak{q}_i}=\sqrt{I}$.
What next?}}
take $y\in
J\smallsetminus\mathfrak{p}_i$. Then $xy^n\in\mathfrak{q}_i$ so
$x\in\mathfrak{q}_i$ since $\mathfrak{q}_i$ is primary and
$y\notin\mathfrak{p}_i$. Hence, $I^{\langle
  J\rangle}\subset\bigcap_{J\nsubset\mathfrak{p}_i}\mathfrak{q}_i$. $\impliedby$
Conversely, suppose that
$x\in\bigcap_{J\nsubset\mathfrak{p}_i}\mathfrak{q}_i$ then
$x\in\mathfrak{q}_i$ for all $\mathfrak{q}_i\nsupset J$. Take any
$\mathfrak{p}_j$ containing $J$. Then
$\mathfrak{p}_j=\nil(R/\mathfrak{q}_j)^c$ (this is easily seen
from the fact that $\mathfrak{p}_i=\sqrt{\mathfrak{q}_i}$, i.e.,
$\mathfrak{q}_i$ is $\mathfrak{p}_i$-primary and the
correspondence theorem for ideals) so there exists $n_j$ with
$xJ^{n_j}\subset\mathfrak{q}_j$ (since, in the quotient, $\bar J$
is nilpotent). Let $n$ be the maximum of all such $n_j$ then
$xJ^n\mathfrak{q}_i$ for all $i$, i.e,
$x\in(I:J^n)=\bigcap_{i}^m(\mathfrak{q}_i:J^n)$. Thus, $x\in
I^{\langle  J \rangle}$.
\\\\
(b) We will prove that $I^{\langle  J\rangle}$ is precisely the
set of all $x\in R$ such that $x/1$ vanishes in
$R_{\mathfrak{p}}$ for all $\mathfrak{p}\nsupset J$. $\implies$
Given $x\in I^{\langle J\rangle}$, $xJ^n\subset I$ for some
$n\geq 1$. Let $\mathfrak{p}$ be a prime ideal not containing $J$
and let $y\in J\smallsetminus\mathfrak{p}$. Then $xy^n\in I$ and
$y^n\notin\mathfrak{p}$ so $x/1=0$ in
$R_{\mathfrak{p}}$. $\impliedby$ Conversely, suppose that $x/1$
vanishes in $R_{\mathfrak{p}}$ for some prime ideal
$\mathfrak{p}\subset R$. Then $xy=0$ for some
$y\in R\smallsetminus\mathfrak{p}$. Since
$\mathfrak{p}=\sqrt{\mathfrak{q}_i}$ for some $i$,
$y^n\in\mathfrak{q}_i$ for some $n\geq 1$. Let
$\bigcap_{i=1}^m\mathfrak{q}_i$ be a minimal decomposition of $0$
(one exists since $R$ is Noetherian) where $\mathfrak{q}_i$ is
$\mathfrak{p}_i$-primary. By part (a), it suffices to show that
\end{proof}
\newpage
\begin{problem}
Let $R$ be a Noetherian ring. Show that $R$ is reduced if and
only if $\Quot(R)$ is a finite direct product of fields.
\end{problem}
\begin{proof}
$\implies$ Suppose that $R$ is reduced. Then $\sqrt{0}=0$. Since
$R$ is Noetherian, by 5.13 every ideal has a primary
decomposition. Let $\bigcap_{i=1}^n\mathfrak{q_i}$, where
$\mathfrak{q}_i$ is $\mathfrak{p}_i$-primary, be such a
decomposition for $0$. Then we have that
$0=\sqrt{0}=\sqrt{\bigcap_{i=1}^n\mathfrak{q}_i}=\bigcap_{i=1}^n\sqrt{\mathfrak{q}_i}=\bigcap_{i=1}^n\mathfrak{p}_i$
so that $\mathfrak{q}_i=\mathfrak{p}_i$. In particular,
$\mathfrak{p}_i$ are minimal over $0$ so by 5.15 the
decomposition is unique. Thus, by the Chinese remainder theorem
we have an exact sequence
\[
0
\longrightarrow
R
\xhookrightarrow{\;\phi\;}
\prod_{i=1}^n R/\mathfrak{p}_i
\xtwoheadrightarrow{\;\pi\;}
\coker\phi
\longrightarrow
0.
\]
Now, consider
$\Quot(R)=S^{-1}R$ where
$S=R\smallsetminus\bigcup_{i=1}^n\mathfrak{p}_i$. By 4.6, since
$S^{-1}R$ is a flat $R$-module, we have an exact sequence
\[
0
\longrightarrow
S^{-1}R
\xhookrightarrow{\;\phi\;}
S^{-1}\left(\prod_{i=1}^nR/\mathfrak{p}_i\right)
\cong
\prod_{i=1}^nS^{-1}R/S^{-1}\mathfrak{p}_i
\xtwoheadrightarrow{\;\pi\;}
S^{-1}\coker\phi
\longrightarrow
0
.
\]
In fact, $\phi$ is an $R$-linear isomorphism since $\coker\phi$
is annihilated by an element of
$S$.\footnote{\textcolor{Red}{Why? We need to show that the set
$\ann_R(\coker\phi)$ is nonempty.}}
It follows that $\Quot(R)\cong \prod_{i=1}^n k_1$ where
$k_i=\Quot(R_{\mathfrak{p}_i})$.

$\impliedby$ Conversely, suppose that $\Quot(R)\cong\prod_{i=1}^n
k_i$ where $k_i$ is a field.
\end{proof}
\newpage
\begin{problem}
Let $R$ be a Noetherian ring and $x\in R$ an $R$-regular
element. Show that $\Ass_R\left(R/(x^n)\right)=\Ass_R(R/(x))$ for
every $n\geq 1$.
\end{problem}
\begin{proof}
Suppose that $x$ is an $R$-regular element then $x\notin\nil(R)$,
i.e., $x$ is not a nilpotent element. Recall that the associated
of $(x^n)$ are precisely $\Ass_R(R/(x^n))$. We will show double
inclusion: One direction is easy namely $\implies$ suppose that
$\mathfrak{p}\in\Ass_R(R/(x^n))$. Then $\mathfrak{p}=\ann(\bar
y)=((x^n):y)$ for some $y\in R$. Then
$\sqrt{\mathfrak{p}}=\sqrt{((x^n):y)}=\bigl(\sqrt{(x^n)}:y\bigr)=((x):y)$
so $\mathfrak{p}\in\Ass_R(R/(x))$. $\impliedby$ Conversely, suppose that
$\mathfrak{p}$ an associated prime of $(x)$.
\end{proof}
\newpage
\begin{problem}
Let $\phi\colon R\to T$ be a homomorphism of rings where $T$ is
Noetherian, let $\prescript{a}{}\phi$ be the induced map on the spectra,
and let $N$ be a $T$-module. Show:
\begin{enumerate}[label=(\alph*)]
\item $\Ass_R(N)=\prescript{a}{}\phi(\Ass_T(N))$.
\item If $N$ is finitely generated as a $T$-module then
  $\Ass_R(N)$ is finite.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $K$ be a field that is a finitely generated
$\ZZ$-algebra. Show that $K$ is a finite field.
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $k$ be a Noetherian ring, $R$ a finitely generated
$k$-algebra, and $\Aut_k(R)$ the group of $k$-algebra
automorphisms of $R$. For a subgroup $G$ of $\Aut_k(R)$ write
$R^G=\left\{\,x\in R\;\middle|\;\text{$\sigma(x)=x$ for every
    $\sigma\in G$}\,\right\}$, which is called the \emph{ring of
  invariants} of $G$. Show that if $G$ is finite then $R^G$ is a
finitely generated $k$-algebra (and hence a Noetherian ring).
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA557-HW-Current"
%%% End:
