\begin{problem}
Let $\varphi\colon R\to S$ be a homomorphism of rings,
$\prescript{a}{}{\varphi}\colon \Spec S\to\Spec R$ the induced map of the
spectra, and $\mathfrak{p}\in\Spec R$. Show that the fiber
$(\prescript{a}{}{\varphi})^{-1}(\mathfrak{p})$ is homeomorphic to
$\Spec(S\otimes_R k(\mathfrak{p}))$.
\end{problem}
\begin{proof}
This is demonstrated (to some extent) by Matsumura, following Theorem 7.2
on p.\,47, we shall attempt to supply the missing details
here. Recall, from the definition of the pre-image, that
\[
(\prescript{a}{}{\varphi})^{-1}(\mathfrak{p})=
\left\{\,\mathfrak{q}\in\Spec S\;\middle|\;\,\mathfrak{q}\cap R=\mathfrak{p}\right\}.
\]
Define a ring homomorphism $\psi\colon S\to S\otimes
k(\mathfrak{p})$ via $\psi(s)\coloneqq s\otimes 1$. We claim that
the induced map on the spectra, i.e., the map
$\prescript{a}{}{\psi}\colon\Spec(S\otimes
k(\mathfrak{p}))\to\Spec S$, is a homeomorphism onto its image
and that
$\im\prescript{a}{}{\psi}=(\prescript{a}{}{\varphi})^{-1}$. First
we will show the latter, that is,
\begin{align*}
\im\prescript{a}{}{\psi}
&=\left\{\,\mathfrak{q}\in\Spec
  S\;\middle|\;\text{$\mathfrak{q}=\mathfrak{r}\cap S$ for
  $\mathfrak{r}\in\Spec(S\otimes k(\mathfrak{p}))$}\,\right\}\\
&=
\end{align*}
\end{proof}
\newpage
\begin{problem}
Let $R\subset S$ be an integral extension of rings with $S$ a Noetherian
ring, and let $\mathfrak{p}\in\Spec R$. Show that there are only finitely
many primes in $S$ lying over $\mathfrak{p}$.
\end{problem}
\begin{proof}
This is the same as Exercise 9.3 from Matsumura, where he
suggests the following approach: Taking a prime ideal
$\mathfrak{p}\in\Spec R$ and localizing at $\mathfrak{p}$ we may
assume that $\mathfrak{p}$ is maximal.
\end{proof}
\newpage
\begin{problem}
Let $\varphi\colon R\to S$ be a homeomorphism of rings with $S$ a
Noetherian ring. Show that the following are equivalent:
\begin{enumerate}[label=(\roman*)]
\item $\varphi$ satisfies going up.
\item $\prescript{a}{}\varphi\colon\Spec S\to\Spec R$ is a closed map.
\item for every $\mathfrak{q}\in\Spec S$, the induced map
  $\Spec(S/\mathfrak{q})\to\Spec(R/\mathfrak{q}\cap S)$ is surjective.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $R\subset S$ be an integral extension of domains with $R$ normal,
$K=\Quot R$,  $\alpha\in S$, $X^n+a_1X^{n-1}+\cdots+a_n$ the minimal
polynomial of $\alpha$ over $K$ (recall $a_i\in R$). Show that for any
$R$-ideal $I$, $\alpha\in \sqrt{IS}$ if and only if $a_i\in\sqrt{I}$ for
$1\leq i\leq n$.
\end{problem}
\begin{proof}
\end{proof}
\newpage
\begin{problem}
Let $k$ be a field and $R=k[X_1,...,X_n]$ a $k$-algebra. Show that the
following are equivalent:
\begin{enumerate}[label=(\roman*)]
\item $R$ is a domain with $\dim R=n-1$
\item $R\cong k[X_1,...,X_n]/(f)$, where $k[X_1,...,X_n]$ is a polynomial
  ring and $f$ is an irreducible polynomial.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA557-HW-Current"
%%% End:
