\section{Product topology}
The aim of this handout is to address two points: metrizability of finite
products of metric spaces, and the abstract characterization of the product
topology in terms of the universal mapping properties among topological
spaces. This latter issue is related to explaining why the definition of
the product topology is not merely \emph{ad hoc} but in a sense the
``right'' definition. In particular, when you study topology more
systematically and encounter the problem of topologizing infinite products
of topological spaces, if you think in terms of the universal property to
be discussed below then you will be inexorably led to the right definition
of the product topology for a product of infinitely many topological spaces
(it is not what one would naively expect to be, based on experience with
the case of finite products).

\subsection{Metrics of finite products}
Let \(X_1,\dotsc,X_d\) be metrizable topological spaces. The product set
\[
  X=\prod_{i=1}^d X_i
\]
admits a natural product topology. It is natural to ask if, upon choosing
metrics \(\rho_j\) inducing the given topology on each \(X_j\), we can
define a metric \(\rho\) on \(X\) in terms of \(\rho_j\) such that \(\rho\)
induces the product topology on \(X\). The basic idea is to find a metric
which describes the idea of ``coordinate-wise closedness'', but several
natural candidates leap out, none of which are evidently better than others
\begin{align*}
  \rho^{\text{max}}\bigl((x_1,\dotsc,x_d),(x_1',\dotsc,x_d')\bigr)
  &=\max_{\mathclap{1\leq j\leq d}}\rho_j(x_j,x_j')\\
  \rho^{\text{Euc}}\bigl((x_1,\dotsc,x_d),(x_1',\dotsc,x_d')\bigr)
  &=\sqrt{\sum_{i=1}^d\rho_j(x_j,x_j')^2}\\
  \rho_1\bigl((x_1,\dotsc,x_d),(x_1',\dotsc,x_d')\bigr)
  &=\sum_{i=1}^d\rho_j(x_j,x_j')\\
  \rho_p\bigl((x_1,\dotsc,x_d),(x_1',\dotsc,x_d')\bigr)
  &=\left[\sum_{j=1}^d\rho_j(x_j,x_j')^p\right]^{\mathrlap{\frac{1}{p}}},&p\geq 1.
\end{align*}

When \(X_j=\bbR\) for all \(j\), with \(\rho_j\) the usual absolute value
metric, these recover the various concrete norms we've seen on
\(X=\bbR^d\). Our first aim will be to show that all of these rather
different-looking metrics are at least bounded above and below by a
positive multiple of each other (which is the best we can expect, since
they sure aren't literally the same), and so in particular they all define
the same topology. In fact, we will see that the common topology they
define is the product topology.

We first axiomatize the preceding examples. Let \(N\colon\bbR^d\to\bbR\) be
any norm which satisfies the property that on the orthant \([0,\infty)^d\)
with nonnegative coordinates it is monotonically increasing function in
each individual coordinate when all others are held fixed. Examples such as
\(N\) include old friends
\[
  \|\blank\|_{\text{max}},\qquad
  \|\blank\|_{\text{Euc}},\qquad
  \|\blank\|_1,\qquad
  \|\blank\|_p,\quad p\geq 1
\]
where we recall that
\[
\|(a_1,\dotsc,a_n)\|_p=\left[\sum_{j=1}^d {|a_j|}^p\right]^{\mathrlap{\frac{1}{p}}}.
\]

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../Diffgeo"
%%% End:
