\subsection{Homework 2}
\begin{problem}[Handout 2, \# 5]
  Four men throw their watches into the sea, and the sea brings each man
  one watch back at random. What is the probability that at least one man
  gets his own watch back?
\end{problem}
\begin{solution*}
  Suppose four men throw their watches into the sea. Place these in some
  order and label them (from left to right) ``the \(k\)\textsup{th} man'',
  \(k=1,\dotsc,4\). Let \(A_k\) denote the event ``the \(k\)\textsup{th}
  man gets his own watch back.'' Then the event \(A\) that at least one man
  gets his own watch back is the union of these events; i.e.,
  \(A=A_1\cup A_2\cup A_3\cup A_4\). Thus, by the inclusion-exclusion
  principle, we have
  \begin{align*}
    P(A)
    &=P(A_1)+P(A_2)+P(A_3)+P(A_4)\\
    &\phantom{{}={}}-P(A_1\cap A_2)-P(A_1\cap A_3)-P(A_1\cap A_4)\\
    &\phantom{{}={}}-P(A_2\cap A_3)-P(A_2\cap A_4)-P(A_3\cap A_4)\\
    &\phantom{{}={}}+P(A_1\cap A_2\cap A_3)+P(A_1\cap A_2\cap A_4)\\
    &\phantom{{}={}}+P(A_1\cap A_3\cap A_4)+P(A_2\cap A_3\cap A_4)\\
    &\phantom{{}={}}-P(A_1\cap A_2\cap A_3\cap A_4).
  \end{align*}
  Since \(P(A_k)=P(A_j)\), \(P(A_k\cap A_j)=P(A_k\cap A_\ell)\), etc., for
  \(k\), \(j\), and \(\ell\) distinct, the equation above reduces to
  \[
    P(A)=4P(A_1)-6P(A_1\cap A_2)+4P(A_1\cap A_2\cap A_3)-P(A_1\cap A_2\cap
    A_3\cap A_4);
  \]
  the choice of \(A_1\), \(A_1\cap A_2\), \(A_1\cap A_2\cap A_3\), etc.\@,
  above was arbitrary.

  All we need to do now is fill in the blanks. The probability that the
  \(1\)\textsup{st} man gets back his own wallet is \(\frac{1}{4}\) since
  only one wallet is his own out of the \(4\). Now, the probability that
  the \(1\)\textsup{st} and the \(2\)\textsup{nd} man get their own wallet
  back is \(\frac{1}{4}\cdot\frac{1}{3}=\frac{1}{2}\). Proceeding in this
  fashion, we have
  \[
    P(A)=4\cdot\tfrac{1}{4}-6\cdot\tfrac{1}{12}+4\cdot\tfrac{1}{24}-\tfrac{1}{24}
    =0.625.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 7]
  Calculate the probability that in Bridge, the hand of at least one player
  is void in a particular suit.
\end{problem}
\begin{solution*}
  Order the players and label them ``player \(k\)'', where
  \(k=1,\dotsc,4\). Let \(A_k\) denote the event ``player \(k\) is void in
  a prespecified suit.'' Then the event \(A\) ``at least one player is void
  in a prespecified suit'' is the union of these events. By the
  inclusion-exclusion principle (and because of the symmetry of these
  evenets) we can decompose the computation of \(P(A)\) to
  \[
    P(A)=4P(A_1)-6P(A_1\cap A_2)+4P(A_1\cap A_2\cap A_3)-P(A_1\cap A_2\cap
    A_3\cap A_4).
  \]

  Let us now fill in the blanks in the equation above. The probability that
  player \(1\) is void in a particular suit, say red clubs
  \(\color{Red}\clubsuit\), is
  \[
    P(A_1)=\frac{\binom{52-13}{13}}{\binom{52}{13}}.
  \]
  Similarly, the probability that player \(1\) and player \(2\) is void in
  \(\color{Red}\clubsuit\) is
  \[
    P(A_1)=\frac{\binom{52-13}{13}}{\binom{52}{13}}%
    \cdot\frac{\binom{52-13-13}{13}}%
    {\binom{52-13}{13}}%
    =\frac{\binom{52-13-13}{13}}{\binom{52}{13}};
  \]
  and so on.

  Thus,
  \[
    P(A)=4\cdot\frac{\binom{52-13}{13}}{\binom{52}{13}}
    -6\cdot\frac{\binom{52-13-13}{13}}{\binom{52}{13}}
    +4\cdot\frac{\binom{52-13-13-13}{13}}{\binom{52}{13}} -1\cdot0\approx
    \num{0.051065521}.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 12]
  If \(n\) balls are placed at random into \(n\) cells, find the
  probability that exactly one cell remains empty.
\end{problem}
\begin{solution*}
  There are \(n\) ways to choose a cell to be left empty. Of the remaining
  \(n-1\) cells, one must contain \(2\) balls. That cell can be chosen in
  \(n-1\) ways and the two balls to be placed in it can be chosen in
  \(\binom{2}{n}\) ways. Of the remaining cells, each must contain \(1\)
  ball and this pairing can be done in \((n-2)!\) ways. Thus, the
  probability of the event \(A\) that exactly one cell remains empty is
  \[
    P(A)=\frac{n(n-1)\binom{2}{n}(n-2)!}{n^n}=\frac{\binom{2}{n}n!}{n^n}.
    \qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 13]
  \emph{(Spread of rumors).} In a town of \(n+1\) inhabitants, a person
  tells a rumor to a second person, who in turn repeats it to a third
  person, etc. At each step the recipient of the rumor is chosen at random
  from the \(n\) people available. Find the probability that the rumor told
  \(r\) times without:
  \begin{enumerate}[label=(\alph*),noitemsep]
  \item returning to the originator,
  \item being repeated to any person.
  \end{enumerate}
  Do the same problem when at each step the rumor told by one person to a
  gathering of \(N\) randomly chosen people. (The first question is the
  special case \(N=1\)).
\end{problem}
\begin{solution*}
  For part (a): The originator can tell any one of the other \(n\)
  inhabitants the rumor. In turn, the non-originator can tell the other
  \(n-1\) inhabitants (not including the originator). Thus, the probability
  of the event \(A\) that the rumor told \(r\) times does not return to the
  originator is
  \[
    P(A)=\frac{n(n-1)^{r-1}}{n^r}=\left(\frac{n-1}{n}\right)^{r-1}.
  \]

  For part (b): The probability of \(B\) that the rumor told \(r\) times is
  not repeated to any person is
  \[
    P(B)=\frac{n(n-1)\dotsc(n-r+1)}{n^r}=\frac{(n)_r}{n^r};
  \]
  the originator is allowed to tell the rumor to any of the other \(n\)
  people while at the \(k\)\textsup{th} step, the person telling the rumor
  is only allowed to tell the rumor to \(n-k\) people.
  \\\\
  Let us calculate part (a) and (b) now for a rumor told to a gathering of
  \(N\) people.

  For part (a): The originator can tell the rumor to any group of \(N\)
  people which can be done in \(\binom{n}{N}\) ways while at the
  \(k\)\textsup{th} step, the person can tell the rumor to any group of
  \(N\) people not including the originator; this can be done in
  \(\binom{n-1}{N}\) ways. Therefore, the probability of \(A_N\) that the
  rumor does not return to the originator if it is told to a gathering of
  \(N\) people is
  \[
    P(A_N)=%
    \frac{\binom{n}{N}\binom{n-1}{N}^{r-1}}{\binom{n}{N}^r}=%
    \left(\frac{\binom{n-1}{N}}{\binom{n}{N}}\right)^{r-1}
  \]

  For part (b): The originator can tell the rumor to any group of \(N\)
  people; this can be done in \(\binom{n}{N}\) ways. At the
  \(k\)\textsup{th} step, the person can tell the rumor to any group of
  \(N\) people not including the previous \(k-1\) people; this can be done
  in \(\binom{n-k+1}{N}\) ways. Therefore, the probability of \(B_N\) that
  the rumor is not repeated to anybody who has already heard it is if it is
  told to a gathering of \(N\) people is
  \[
    P(B_N)=\frac{\binom{n}{N}\dotsm\binom{n-r+1}{N}}{\binom{n}{N}^r}.
  \]
  (This equation breaks down for \(n-r+1<N\).)
\end{solution*}

\begin{problem}[Handout 2, \# 14]
  What is the probability that
  \begin{enumerate}[label=(\alph*),noitemsep]
  \item the birthdays of twelve people will fall in twelve different
    calendar months (assume equal probabilities for the twelve months),
  \item the birthdays of six people will fall in exactly two calendar
    months?
  \end{enumerate}
\end{problem}
\begin{solution*}
  For part (a): Let \(A\) denote the event that the birthdays of twelve
  people will fall in twelve different calendar months. There are \(12!\)
  ways to assign a person to a calendar date without
  duplication. Therefore,
  \[
    P(A)=\frac{12!}{12^{12}}= \frac{\num{479001600}}{\num{8916100448256}}
    \approx\num{5.372321709247828e-05}
  \]

  For part (b): Let \(B\) denote the event that the birthdays of six people
  will fall in exactly two calendar months. There are \(\binom{12}{2}=66\)
  ways to choose two out of twelve calendar months. Now, we can have one
  person having his birthday on the

  For part (b): Let \(B\) denote the event that the birthdays of six people
  will fall in exactly two calendar months. First, there are
  \(\binom{12}{2}=66\) ways to choose the two calendar months on which the
  six people's birthdays will fall. For each of the two months, we can have
  one person having his birthday on the first and eleven on the other; two
  persons having their birthday on the first and ten on the other; etc.\@
  and vice-versa. Therefore,
  \[
    P(B)=%
    \frac{\binom{12}{2}\cdot \binom{2}{1}%
      \cdot\left[\binom{6}{1}+\dotsb+\binom{6}{5}\right]}%
    {12^6}%
    \approx \num{2.7408050411522633e-03}.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 15]
  A car is parked among \(N\) cars in a row, not at either end. On his
  return the owner finds exactly \(r\) of the \(N\) places still
  occupied. What is the probability that both neighboring places are empty?
\end{problem}
\begin{solution*}
  Let \(A\) denote the event that upon returning the car owner finds the
  parking spots neighboring his car empty. Let us count all of the possible
  arrangements that leave the two spots adjacent to the car owner's car
  empty. There were originally \(N-1\) cars (excluding the car owner's car)
  so there are \(\binom{N-1}{r-1}\) possible arrangements for the remaining
  cars (excluding the car owner's car); i.e.,
  \(|\Omega|=\binom{N-1}{r-1}\). Now, since we want to count those
  arrangements that leave the two spots adjacent to the car owner's car
  empty, two of the choices above are forced on us. Therefore, the number
  of arrangements which leave the two spots adjacent to the car owner's car
  empty is \(\binom{N-1-2}{r-1}\). Thus,
  \[
    P(A)=\frac{\binom{N-3}{r-1}}{\binom{N-1}{r-1}}.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 16]
  Find the probability that in a random arrangement of \(52\) bridge card
  no two aces are adjacent.
\end{problem}
\begin{solution*}
  Let \(A\) denote the event ``there are no two aces adjacent in a random
  arrangement of \(52\) bridge card.'' We can arrange \(52\) cards in
  \(52!\) ways so \(|\Omega|=52!\). Now, we can arrange four aces in any of
  \(4!\) ways and, so that we do not place any ace adjacent to one another,
  place them in \(\binom{48+1}{4}=\binom{49}{4}\) slots in the original
  \(52\) set of cards (including the end places). The rest of the \(48\)
  cards can be arranged in \(48!\) ways. Thus,
  \[
    P(A)=\frac{\binom{49}{4}4!48!}{52!}\approx\num{0.7826244343891402}.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 17]
  Suppose \(P(A)=\frac{3}{4}\), and \(P(B)=\frac{1}{3}\).

  \noindent Prove that \(P(A\cap B)\geq\frac{1}{12}\). Can it be equal to
  \(\frac{1}{12}\)?
\end{problem}
\begin{solution*}
  By the inclusion-exclusion principle, we have
  \[
    P(A\cup B)=P(A)+P(B)-P(A\cap B)
  \]
  so
  \[
    P(A\cap B)=P(A)+P(B)-P(A\cup B).
  \]
  But, since \(A\cup B\subset\Omega\) and \(P(\Omega)=1\), \(P(A\cup
  B)\leq P(\Omega)=1\) so \(-P(A\cup B)\geq -1\). Thus,
  \begin{align*}
    P(A\cap B)&=P(A)+P(B)-P(A\cup B)\\
              &\geq P(A)+P(B)-1\\
              &=\frac{3}{4}+\frac{1}{3}-1\\
              &=\frac{9}{12}+\frac{4}{12}-\frac{12}{12}\\
              &=\frac{1}{12}.
  \end{align*}

  Lastly, we show that \(P(A\cap B)\) can in fact be equal to
  \(\frac{1}{12}\). Consider the interval \(\Omega\defeq [0,1]\) equipped
  with the probability measure \(P(I)=b-a\) where \(I\) is any interval
  with endpoints \(\{a,b\}\) contained in \(\Omega\). Set
  \(A\defeq (0,\frac{3}{4})\) and
  \(B\defeq (1-\frac{1}{3},1)=(\frac{2}{3},1)\). Then the intersection
  \(A\cap B=(\frac{2}{3},\frac{3}{4})\) has probability
  \[
    P(A\cap B)=\frac{3}{4}-\frac{2}{3}=\frac{1}{12}.\qedhere
  \]
\end{solution*}

\begin{problem}[Handout 2, \# 18]
  Suppose you have infinitely many events \(A_1,A_2,\dotsc\), and each one
  is sure to occur, i.e., \(P(A_k)=1\) for any \(i\).

  \noindent Prove that \(P\bigl(\bigcap_{k=1}^n A_k\bigr)=1\).
\end{problem}
\begin{solution*}
  Consider the sequence of probabilities \(\{P_n\}\) where
  \(P_n=P\bigl(\bigcap_{i=1}^n A_i\bigr)\). Note that
  \(\bigcap_{i=1}^n A_i\downarrow\bigcap_{i=1}^\infty A_i\). First we show,
  by induction, that \(P_n=1\).

  The case \(n=1\) is trivial. Now, assume the result holds for \(n-1\) and
  consider \(P_n=P\bigl(\bigcap_{i=1}^n A_i\bigr)\). Writing
  \(A'=\bigcap_{i=1}^{n-1}A_i\), we have
  \[
    P_n=P(A'\cap A_n).
  \]
  By the inclusion-exclusion principle,
  \begin{align*}
    P_n&=P(A')+P(A_n)-P(A'\cup A_n)\\
       &=1+1-P(A'\cup A_n)
         \intertext{and since \(P(A'\cup A_n)\geq P(A')=1\) by the monotonicity
         of the probability measure, \(P(A'\cup A_n)=1\) since \(P(A'\cup
         A_n)\leq P(\Omega)=1\), thus}
       &=1+1-1\\
       &=1.
  \end{align*}
  It follows that \(\{P_n\}\) is the constant sequence \(\{1\}\) therefore,
  its limit is \(1\).
\end{solution*}

\begin{problem}[Handout 2, \# 19]
  There are \(n\) blue, \(n\) green, \(n\) red, and \(n\) white balls in an
  urn. Four balls are drawn from the urn with replacement. Find the
  probability that there are balls of at least three different colors among
  the four drawn.
\end{problem}
\begin{solution*}
  Let \(A\) denote the event ``there are balls of at least three different
  colors among the four drawn.'' There are \(4^4\) ways to draw four balls;
  therefore, \(|\Omega|=4^4\). There are \(4!\) wasy to draw a ball of each
  color and \(4\cdot 3\binom{4}{2} 2!\) ways to draw four balls missing
  exactly one color (pic a color to be missed; pick a color to be drawn
  twice; pick two draws for that color to be drawn on; and rearrange the
  other two colors into the other two draws). Thus,
  \[
    P(A)=\frac{4!+4\cdot 3\binom{4}{2}2!}{4^4}\approx\num{0.65625}.\qedhere
  \]
\end{solution*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA519-HW-ALL"
%%% End:
