\begin{problem}[Handout 10, \# 4]
  \emph{(Poisson Approximation.)} One hundred people will each toss a fair
  coin \(200\) times. Approximate the probability that at least \(10\) of
  the \(100\) people would each have obtained exactly \(100\) heads and
  \(100\) tails.
\end{problem}
\begin{solution}
  Let \(X\) denote the number of people who obtain exactly \(100\) heads
  and (consequently) \(100\) tails. First, we compute the probability that
  any one given person obtains exactly \(100\) heads. There are \(2^{200}\)
  possible outcomes for \(200\) tosses of a fair coin, and
  \(\binom{200}{100}\) possible ways of obtaining exactly \(100\)
  heads. Thus, the probability that any one person obtains exactly \(100\)
  head in \(200\) tosses of a fair coin is
  \[
    p=\frac{\binom{200}{100}}{2^{200}}\approx\num{0.05634847900925642}.
  \]

  Now, assuming \(X\sim\Poisson(\num{5.634847900925642})\), the
  probability that at least \(10\) of the \(100\) people have each obtained
  exactly \(100\) heads and \(100\) tails is
  \begin{align*}
    P(X\geq 10)
    &=1-P(X<10)\\
    &=1-\sum_{i=1}^9 P(X=i)\\
    &=1
      -e^{-\num{5.634847900925642}}
      \sum_{i=0}^9\frac{\num{5.634847900925642}^i}{i!}\\
    &=1-\num{0.8825634032515405}\\
    &\approx \num{0.11743659674845952}.
  \end{align*}
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 5]
  \emph{(A Pretty Question.)} Suppose \(X\) is a Poisson distributed random
  variable. Can three different values of \(X\) have an equal probability?
\end{problem}
\begin{solution}
  No. Let \(X\sim\Poisson(\lambda)\). First, we show that given any two
  values \(k_1,k_2\in\bbZ_{\geq 0}\), there exists \(\lambda\) such that
  \(p(k_1)=p(k_2)\).

  Observe that for \(p(k_1)=p(k_2)\) we must have
  \begin{align*}
    e^{-\lambda}\frac{\lambda^{k_1}}{k_1!}
    &=e^{-\lambda}\frac{\lambda^{k_2}}{k_2!}\\
    \lambda^{k_1-k_2}
    &=\frac{k_1!}{k_2!}
      \intertext{this implies that, given \(k_1\) and \(k_2\),}
      (k_1-k_2)\ln\lambda&=\ln(k_1!/k_2!)\\
    \lambda(k_1,k_2)&=e^{\ln(k_1!/k_2!)/(k_1-k_2)}.
  \end{align*}
  For example, \(\lambda(3,5)\approx\num{4.472135954999579}\) and
  \[
    p(3)\approx\num{0.17028240507308132}\approx p(5).
  \]

  We now prove our original claim. Assume for a moment that \(X\) is
  continuous. We will show that the PMF \(p\) of \(X\) has at most one
  critical point. Write the PMF of \(p\) as its continuous analogue
  \[
    p(x)=P(X=x)=\frac{e^{-\lambda}\lambda^x}{\Gamma(x)}.
  \]
  Then, taking the derivative of \(p\), we have
  \begin{align*}
    p'(x)
    &=e^{-\lambda}\frac{\Gamma(x)\lambda^x\ln\lambda-\lambda^x\Gamma(x)}{\Gamma(x)^2}\\
    &=\frac{e^{-\lambda}\lambda^x}{\Gamma(x)}
      \left[\ln\lambda-\frac{\Gamma'(x)}{\Gamma(x)}\right]\\
    &=p(x)\left[\ln\lambda-\frac{\Gamma'(x)}{\Gamma(x)}\right].
  \end{align*}
  Since \(\Gamma,\Gamma'>0\) for all \(x\in\bbR_{\geq 0}\), \(p'(x)=0\) if
  and only if
  \[
    \ln\lambda=\frac{\Gamma'(x)}{\Gamma(x)}
  \]
  which happens at most once since the quotient.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 6]
  \emph{(Poisson Approximation.)} There are \(20\) couples seated at a
  rectangular table, husbands on one side and the wives on the other, in a
  random order. Using a Poisson approximation, find the probability that
  exactly two husbands are seated directly across from their wives; at
  least three are; at most three are.
\end{problem}
\begin{solution}
  Let \(X\) count the number of husbands that have been seated directly
  across from their respective wives. We approximate the probabilities that
  (i) exactly two husbands are seated directly across from their wives,
  (ii) at least three are, and (iii) at most three are, by assuming that
  \(X\sim\Poisson(\lambda)\). But first, we compute the mean of \(X\).

  Fixing the men, the probability that any given man is sitting directly
  across from his wife is \(1/20\). Since the distribution of \(X\) is
  binomial, its mean is \(20(1/20)=1\). Thus, \(\lambda=1\).

  For (i), the probability that exactly two husbands are seated directly
  from their wives is
  \[
    p(2;1)=e^{-1}\frac{1}{2!}\approx\num{0.18393972058572117}.
  \]

  For (iii), the probability is
  \[
    p(x\geq
    3;1)=1-e^{-1}\left[1+1+\frac{1}{2}\right]\approx\num{0.08030139707139416}.
  \]

  For (iii), the probability is
  \[
    p(x\leq 3;1)=
    e^{-1}\left[1+1+\frac{1}{2}+\frac{1}{6}\right]\approx\num{0.9810118431238462}.
  \]
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 7]
  \emph{(Poisson Approximation.)} There are \(5\) coins on a desk, with
  probabilities \(0.05\), \(0.1\), \(0.05\), \(0.01\), and \(0.04\) for
  heads. By using a Poisson approximation, find the probability of
  obtaining at least one head when the five coins are each tossed once.

  \noindent Is the number of heads obtained binomially distributed in this
  problem?
\end{problem}
\begin{solution}
  Let \(X_i\), \(1\leq i\leq 5\), count the number of heads of a toss of
  the \(i\)\textsup{th} coin. First, let us find the mean of
  \(X=X_1+\dotsb+X_5\). This is given by
  \[
   0.05+0.1+0.05+0.01+0.04=0.25.
  \]
  Set \(\lambda=0.25\). Then, using a Poisson approximation to the PMF of
  \(X\), the probability of obtaining at least one head is
  \[
    e^{-0.25}\sum_{i=1}^5 \frac{0.25^i}{i!}\approx\num{0.22119894311503177}.
  \]

  Theorem 1.28(a) from the Text suggests that the distribution may not be
  binomial unless the probabilities of each head are identical.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 8]
  A book of \(500\) pages contains \(500\) misprints. Estimate the chances
  that a given page contains at least three misprints.
\end{problem}
\begin{solution}
  We approximate the distribution of \(X\) the number of misprints on a
  given page using a Poisson distribution with parameter \(\lambda=1\) the
  density of misprints throughout the book. Under these assumptions, the
  probability that a given page contains at least three misprints is given
  by
  \[
    1-e^{-1}\frac{1}{1!}-e^{-1}\frac{1}{1!}-e^{-1}\frac{1}{2!}
    \approx\num{0.08030139707139416}.
  \]
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 9]
  Estimate the number of raisins which a cookie should contain on the
  average if it is desired that not more than one cookie out of a hundred
  should be without raisin.
\end{problem}
\begin{solution}
  The number of raisins in a cookie is given by a poisson distribution: Let
  $X_{\lambda}$ be the number of raisins in a cookie that came out of a batch
  where the average number of raisins per cookie was $\lambda$. We want
  $P(X_{\lambda} = 0) \leq 0.01$. That is, we want
  \[
    \frac{\lambda^0 e^{-\lambda}}{0!} = e^{-\lambda} \leq 0.01
  \]
  which occurs for
  $\lambda \geq -\ln(0.01) = \ln(100) \approx \num{4.60517}$.

  That is, we want there to be about $4.6$ raisins per cookie in order to
  guarantee that no more than one out of one hundred cookies is raisinless.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 10]
  The terms \(p(k;\lambda)\) of the Poisson distribution reach their
  maximum when \(k\) is the largest integer not exceeding \(\lambda\).
\end{problem}
\begin{solution}
  First, set $\lfloor{\lambda}\rfloor$ equal to the largest integer not
  exceeding $\lambda$. Then we show $p(k;\lambda)$ is increasing on
  $[0,\lfloor{\lambda}\rfloor]$ and decreasing on
  $[\lfloor{\lambda}\rfloor, \infty)$; this suffices to show that the maximum
  occurs at $\lfloor{\lambda}\rfloor$.

  Now, note that for all $k$,
  \begin{align*}
    p(k; \lambda) - p(k+1;\lambda)
    &= \frac{\lambda^k e^{-\lambda}}{k!} -\frac{\lambda^{k+1} e^{-\lambda}}{(k+1)!}\\
    &= \frac{e^{-\lambda} \lambda^k}{k!} \left[ 1 -\frac{\lambda}{k+1} \right]\\
    &= \frac{e^{-\lambda} \lambda^k}{k!} \left[ 1 -\frac{\lambda}{k+1} \right]
  \end{align*}
  which is positive exactly when $k < \lambda +1$ and negative exactly when
  $k \geq lambda +1$. That is, $p(k+1;\lambda)$ is larger than
  $p(k; \lambda)$ exactly when $k < \lambda +1$; that is, $p(k;\lambda)$
  increases until the largest integer not exceeding $\lambda$;
  $p(k;\lambda)$ is increasing on $[0,\lfloor {\lambda} \rfloor$ and decreasing on
  $[\lfloor  {\lambda}\rfloor,\infty)$, as desired.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 11]
  Prove
  \[
    p(0,\lambda)+\dotsb+p(n,\lambda)
    =\frac{1}{n!}\int_\lambda^\infty e^{-x}x^n\diff x.
  \]
\end{problem}
\begin{solution}
  By integration by parts, we have
  \begin{align*}
    \frac{1}{n!}\int_\lambda^\infty e^{-x}x^n\diff x
    &=\frac{1}{n!}e^{-x}x^n\Bigr|_\lambda^\infty-\frac{1}{n!}\int_\lambda^\infty
      -ne^{-x}x^{n-1}\diff x\\
    &=\frac{e^{-\lambda}\lambda^n}{n!}+\frac{1}{(n-1)!}\int_\lambda^\infty
      e^{-x}x^{n-1}\diff x\\
    &\vdotswithin{=}\\
    &=\frac{e^{-\lambda}\lambda^n}{n!}+\frac{e^{-\lambda}\lambda^{n-1}}{(n-1)!}+\dotsb+
      \frac{e^{-\lambda}\lambda^{0}}{0!}\\
    &=p(n;\lambda)+\dotsb+p(0;\lambda),
  \end{align*}
  as was to be shown.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 12]
  There is a random number \(N\) of coins in your pocket, where \(N\) has a
  Poisson distribution with mean \(\mu\). Each one is tossed once.

  \noindent Let \(X\) be the number of times a head shows.

  \noindent Find the distribution of \(X\).
\end{problem}
\begin{solution}
  First, note that $P(N=n)=e^{-\mu} \mu^n/n!$.  Given that there are $n$
  coins in your pocket, let $X_n$ denote the number of heads shown. Then
  $P(X_n = m) = \binom{n}{m} \left( 0.5 \right)^n$.  So, the distribution
  of $X$ is given by
  \begin{align*}
    P(X=m) &= \sum_{n=0}^\infty P(X_n = m) P(N=n) \\
           &= \sum_{n=0}^\infty \binom{n}{m} \left( 0.5 \right)^n \frac{e^{-\mu} \mu^n}{n!} \\
           &= e^{-\mu} \sum_{n=0}^\infty \frac{n!}{(n-m)!m!n!} (0.5 \mu)^n \\
           &= \frac{e^{-\mu}}{m!} \sum_{n=0}^\infty \frac{(0.5 \mu)^n}{(n-m)!}
  \end{align*}
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 14]
  Find the MGF of a general Poisson distribution, and hence prove that the
  mean and the variance of an arbitrary Poisson distribution are equal.
\end{problem}
\begin{solution}
  Let $X$ be a poisson distribution with mean $\lambda$. Then
  \begin{align*}
    M_X(t) &= \sum_{n=0}^\infty e^{tn} P(X = n) \\
           &= \sum_{n=0}^\infty e^{tn} \frac{e^{-\lambda} \lambda^n}{n!} \\
           &= e^{-\lambda} \sum_{n=0}^\infty \frac{(e^t \lambda)^n}{n!} \\
           &= e^{-\lambda} e^{\lambda e^t} \\
  \end{align*}
  Note that
  $M_X'(t) = e^{-\lambda} \left(\lambda e^{\lambda e^t + t} \right)$ and
  $M_X''(t) = e^{-\lambda} \left(\lambda e^{\lambda e^t + t} \right)
  \left(\lambda e^t +1 \right)$.

  Evaluating at $0$, we get that $E(X) = M_X'(0) = \lambda$ and
  $E(X^2) = M_X''(0) /2= \lambda$.
\end{solution}
\newpage

\begin{problem}[Handout 10, \# 17 (a)]
  \emph{(Poisson approximations.)} \(20\) couples are seated in a
  rectangular table, husbands on one side and the wives on the
  other. First, find the expected number of husbands that sit directly
  across from their wives. Then, using a Poisson approximation, find the
  probability that two do; three do; at most five do.
\end{problem}
\begin{solution}
  Like in Problem 7.3, \(\lambda=1\).

  Using a Poisson approximation, the probability that two husbands are
  seated directly across from their wives is
  \[
    p(2;1)=e^{-1}\frac{1}{2}\approx\num{0.18393972058572117}.
  \]

  The probability that three do is
  \[
    p(3;1)=e^{-1}\frac{1}{6}\approx\num{0.06131324019524039}.
  \]

  The probability that at most five do
  \[
    p(x\leq 5;1)=%
    e^{-1}%
    \left[1+1+\frac{1}{2}+\frac{1}{6}+\frac{1}{24}+\frac{1}{120}\right]\approx%
    \num{0.9994058151824182}.
  \]
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA519-HW-Current"
%%% End:
