\subsubsection{Homework 9}
\setcounter{exercise}{0}
\setcounter{equation}{0}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 1]
  \hfill
  \begin{enumerate}[label=(\alph*),noitemsep]
  \item Let \(E\) be a measurable subset of \(\bbR^2\) such that for almost
    every \(x\in\bbR\), \(\left\{\,y:(x,y)\in E\,\right\}\) has
    \(\bbR\)-measure zero. Show that \(E\) has measure zero and that for
    almost every \(y\in\bbR\), \(\left\{\,x:(x,y)\in E\,\right\}\) has
    measure zero.
  \item Let \(f(x,y)\) be nonnegative and measurable in \(\bbR^2\). Suppose
    that for almost every \(x\in\bbR\), \(f(x,y)\) is finite for almost
    every \(y\). Show that for almost \(y\in\bbR\), \(f(x,y)\) is finite
    for almost every \(x\).
  \end{enumerate}
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 3]
  Let \(f\) be measurable and finite a.e.\@ on \([0,1]\). If \(f(x)-f(y)\)
  is integrable over the square \(0\leq x\leq 1\), \(0\leq y\leq 1\), show
  that \(f\in L[0,1]\).
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 4]
  Let \(f\) be measurable and periodic with period \(1\):
  \(f(t+1)=f(t)\). Suppose there is a finite \(c\) such that
  \[
    \int_0^1|f(a+t)-f(b+t)|\diff t\leq c
  \]
  for all \(a\) and \(b\). Show that \(f\in L[0,1]\). (Set \(a=x\),
  \(b=-x\), integrate with respect to \(x\), and make the change of
  variables \(\xi=x+t\), \(\eta=-x+t\).)
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 6]
  For \(f\in L(\bbR)\), define the \emph{Fourier transform \(\hat f\)} of
  \(f\) by
  \[
    \hat f(x)=\frac{1}{2\pi}\int_{-\infty}^\infty f(t)e^{-ixt}\diff t
  \]
  for \(x\in\bbR\). (For complex-valued function \(F=F_0+iF_1\) whose real
  and imaginary parts \(F_0\) and \(F_1\) are integrable, we define
  \(\int F=\int F_0+i\int F_1\).) Show that if \(f\) and \(g\) belong to
  \(L(\bbR)\), then
  \[
    \widehat{(f*g)}(x)=2\pi\hat f(x)\hat g(x).
  \]
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 7]
  Let \(F\) be a closed subset of \(\bbR\) and let
  \(\delta(x)=\delta(x,F)\) be the corresponding distance function. If
  \(\lambda>0\) and \(f\) is nonnegative and integrable over the complement
  of \(F\), prove that the function
  \[
    \int_{\bbR}\frac{\delta^\lambda(y)f(y)}{|x-y|^{1+\lambda}}\diff t
  \]
  is integrable over \(F\) and so is finite a.e.\@ in \(F\). (In case
  \(f=\chi_{(a,b)}\), this reduces to Theorem 6.17.)
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 6, Ex.\@ 9]
  \hfill
  \begin{enumerate}[label=(\alph*)]
  \item Show that \(M_\lambda(x;F)=+\infty\) if \(x\notin F\),
    \(\lambda>0\).
  \item Let \(F=[c,d]\) be a closed subinterval of a bounded open interval
    \((a,b)\subseteq\bbR\), and let \(M_\alpha\) be the corresponding
    Marcinkiewicz integral, \(\lambda>0\). Show that \(M_\lambda\) is
    finite for every \(x\in(c,d)\) and that
    \(M_\lambda(c)=M_\lambda(d)=\infty\). Show also that
    \(\int M_\lambda\leq\lambda^{-1}|G|\), where \(G=(a,b)-[c,d]\).
  \end{enumerate}
\end{problem}
\begin{solution}
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
