% MA 54400 Homework assignment - Spring 2016
% Problems in parentheses are relative to material not yet covered in class.
% #1 - Due Jan. 20 Read Sections 2.1, 2.3; Chapter 2: # 1, 2, 3, 11, 13.
% #2 - Due Jan. 25 Read Section 3.1; Problem #1: Show that the boundary of
% any interval has outer measure zero; Problem #2: Show that a set
% consisting of a single point has outer measure zero.
% #3 - Due Feb. 1 Read Sections 3.2-4.  Chapter 3: # 5, 7, 8, 9, 10.
% #4 - Due Feb. 8 Read Section 3.5. Chapter 3: # 12, 13, 15.
% #5 - Due Feb. 15 Read Sections 3.6, 4.1 Chapter 3: # 14, 16, 18; Chapter
% 4: # 1, 2.
% Practice Problems for Midterm 1 - Solutions
% #6 - Due Feb. 22 Read Sections 4.2-3. Chapter 4: # 4, 7, 8.
% #7 - Due Mar. 2 Read Sections 5.1-2. Chapter 4: #9, 11, 15, 18 (ignore
% the second part concerning convergence in measure); Chapter 5: # 1, 3.
% #8 - Due Mar. 7 Read Section 5.3, 6.1. Chapter 5: # 2, 4, 6, 7,
% 21. Chapter 6: # 10, 11.
% #9 - Due Mar. 21 Read Section 6.2-3. Chapter 5: # 20; 6: # 1, 3, 4, 6, 7,
% 9.
% Practice Problems for Midterm 2 (You can ignore #4) – Solutions
% #10 – Due Apr. 11 Read Section 7.1-6. Chapter 7: # 1, 2, 12, 13, 6, 8, 9
% #11 – Due Apr. 18 Read Section 5.4, Sections 8.1-4. Chapter 7: # 11, 15;
% Chapter 5: # 8, 11, 12, 17; Prove Theorem 8.3.
% #12 – Due Apr. 25 Chapter 8: # 2, 4, 5, 6, 9, 11, 16, 18.
% Final Exam on Tuesday, May 3, 8-10 am, UNIV 117
% Practice Problems for the Final Exam - Solutions

% Interesting symbols
\section{MA 544 Spring 2016}
% \thispagestyle{empty}
This is material from the course MA 544 as it was taught in the spring of
2016.%
\bigskip
\subsection{Homework}
These exercises were assigned from Wheeden and Zygmund's \emph{Measure and
  Integral}, therefore, most of the theorems I reference will be from
\cite{wheeden-zygmund}. Other resources include \cite{folland} and
\cite{royden}. For more elementary results, I cite \cite{rudin-1}. Unless
otherwise stated, whenever we quote a result, e.g., Theorem 1.1, it is
understood to come from Wheeden and Zygmund's \emph{Measure and Integral}.

\bigskip

Throughout these notes

\begin{tabular}{cl}
  \(\bbR\) & is the set of real numbers\\
  \(\bbR^+\) & is the set of positive real numbers, that is, \(x\in\bbR\) with
             \(x\geq 0\)\\
  \(\bbC\) & is the set of complex numbers\\
  \(\bbQ\) & is the set of rational numbers\\
  \(\bbZ\) & is the set of the integers\\
  \(\bbZ^+\) & is the set of positive integers, that is, \(x\in\bbZ\) with
             \(x\geq 0\)\\
  \(\bbN\) & is the set of the natural numbers \(1,2,\dotsc\)\\
  \(A\setminus B\) & is the set difference of \(A\) and \(B\), that is, the
                        complement of \(A\cap B\) in \(A\)\\
  \(m^*(E)\) & the outer measure of \(E\)\\
  \(m_*(E)\) & the inner measure of \(E\)\\
  \(m(E)\) & the Lebesgue measure of \(E\)\\
  \(\|\blank\|\) & the standard Euclidean norm on \(\bbR^n\)\\
  \(f\asymp g\) & means \(f\) is asymptotically equivalent to \(g\), that is,
                \(\lim_{x\to\infty} g(x)/f(x)=1\)\\
\end{tabular}

\newpage

\subsubsection{Homework 1}
\setcounter{exercise}{0}
\setcounter{equation}{0}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 2, Ex.\@ 1]
  Let \(f(x)=x\sin(1/x)\) for \(0<x\leq 1\) and \(f(0)=0\). Show that \(f\) is
  bounded and continuous on \([0,1]\), but that \(V[f;0,1]=\infty\).
\end{problem}
\begin{solution}
  Let \(f\) equal \(x\sin(1/x)\). We will show that \(f\) is bounded and
  continuous on \([0,1]\), but that it is not of bounded variation on
  \([0,1]\).

  First we will show that \(f\) is bounded. Note that both \(|x|\) and
  \(|\sin(1/x)|\) are bounded by \(1\) on the interval \([0,1]\). Since
  \(|f|=|x||\sin(1/x)|\), it follows that \(|f|\leq 1\) on \([0,1]\). Thus, \(\)f
  is bounded on \([0,1]\).

  Next we show that \(f\) is continuous. It is easy to show that \(f\) is
  continuous on the subinterval \((0,1]\) since both \(|x|\) and \(\sin(1/x\)
  are continuous on that interval and we know that the product of
  continuous functions is continuous. To see that \(f\) is continuous at \(\)0
  we must show that \(f(x^+)=f(0)\); that is, the limit of \(f\) as \(\)x
  approaches \(0\) from the right is \(f(0)\) which by definition is \(0\). To
  this end, it suffices to take a (monotonically decreasing) sequence
  \(x_n\downarrow 0\) and show that the limit of the sequence
  \({\{f(x_n)\}}_{n=1}^\infty\) is \(0\). Let \(\varepsilon>0\) be given then,
  since \(x_n\) converges to \(0\) there exists an index \(N\) such that
  \(|0-x_n|<\varepsilon\) whenever \(n\geq N\). Since \(|f(x_n)|\leq |x_n|\) on
  \([0,1]\), the following inequality holds
  \begin{align*}
    |0-f(x_n)|
    &=\left|0-x_n\sin(1/x_n)\right|\\
    &\leq |x_n|\\
    &<\varepsilon.
  \end{align*}
  Thus, \(f\) is continuous at \(0\) and it converges to \(0\).

  Despite the nice properties that \(f\) seemingly possesses, \(f\) is not
  b.v.\@ on \([0,1]\). To show that \(f\) is not b.v.\@ on \([0,1]\) we
  must show that for any positive real number \(M\) there exists some
  partition \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of \([0,1]\) such that
  the sum associated to \(\Gamma\)
  \[
    \sum_{i=1}^n|f(x_i)-f(x_{i-1})|>M.
  \]
  Let \(N\) be the smallest integer greater than \(M\) and let \(n\) be the
  smallest integer greater than or equal to \(N/2\). Then the partition
  \(\Gamma=\{\,x_0=1<x_1<\cdots<x_{n+1}=1\,\}\) where \(x_i=2/((3+(n-i))\pi\)
  for \(1\leq i\leq N\). Then we have the inequality
  \begin{align*}
    S_\Gamma
    &=\sum_{i=1}^{n+1}|f(x_i)-f(x_{i-1})|\\
    &=\sum_{i=2}^n
      |f(x_i)-f(x_{i-1})|+|f(x_{n+1})-f(x_n)|+|f(x_0)-f(x_1)|\\
    &=N+|f(x_{n+1})-f(x_n)|+|f(x_0)-f(x_1)|\\
    &>M.
  \end{align*}
  Thus, \(f\) is not b.v.\@ on \([0,1]\).
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 2, Ex.\@ 2]
  Prove theorem (2.1).
\end{problem}
\begin{solution}
  Recall the statement of Theorem 2.1:
  \begin{quote}
    \begin{enumerate}[label=(\alph*)]
    \item If \(f\) is of bounded variation on \([a,b]\), then \(f\) is bounded on
      \([a,b]\).
    \item Let \(f\) and \(g\) be of bounded variation on \([a,b]\). Then
      \(cf\) (for any real constant \(c\)), \(f+g\), and \(fg\) are of
      bounded variation on \([a,b]\). Moreover, \(f/g\) is of bounded
      variation on \([a,b]\) if there exists an \(\varepsilon>0\) such that
      \(|g(x)|\geq\varepsilon\) for \(x\in[a,b]\).
    \end{enumerate}
  \end{quote}
  \noindent%
  We shall prove these in alphabetical order:

  For part (a) we shall proceed by contradiction. First, without loss of
  generality, we may assume that \(f(a)=0\) since the function the
  variation of \(g(x)=f(x)-f(a)\) is equal to the variation of \(f\) and
  \(g(a)=0\).  Suppose that \(f\) is b.v.\@ on \([a,b]\) with variation
  \(V=V[f;a,b]\), but that \(f\) is unbounded on \([a,b]\); that is, given
  a positive real number \(M\) there exists a point \(x\) in \([a,b]\) such
  that \(|f(x)|>M\). In particular, there exists \(x\in[a,b]\) such that
  \(|f(x)|>V\). Hence, for any \(x\in[a,b]\) by the triangle inequality we
  have
  \begin{align*}
    V&<|f(x)|\\
     &=|f(x)-f(a)+f(a)|\\
     &\leq|f(x)-f(a)|+|f(a)|\\
     &\leq V.
  \end{align*}
  This is a contradiction. Therefore, it must be the case that if \(f\) is
  b.v.\@ on \([a,b]\) then \(f\) is bounded on \([a,b]\).

  We break part (b) into three sections. Suppose \(f\) and \(g\) are b.v.\@
  on \([a,b]\) with variation \(V\) and \(V'\), respectively.  We will show
  that (i) \(cf\); (ii) \(f+g\); and (iii) \(fg\) are b.v.\@ on
  \([a,b]\). Moreover, we show that (iv) \(f/g\) is b.v.\@ on \([a,b]\) if
  there exists \(\varepsilon>0\) such that \(|g(x)|\geq\varepsilon\) for
  all \(x\in[a,b]\).

  For part (i) above let \(c\) be a real number. Given a partition
  \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of \([a,b]\), we have
  \begin{align*}
    S_\Gamma
    &=\sum_{i=1}^n |cf(x_i)-cf(x_{i-1})|\\
    &=\sum_{i=1}^n |c||f(x_i)-cf(x_{i-1})|\\
    &=|c|\sum_{i=1}^n|f(x_i)-f(x_{i-1})|\\
    &\leq |c| V
  \end{align*}
  since \(V\) is the supremum of the sums of the form
  \(\sum_{i=1}^m|f(x_i)-f(x_{i-1})|\) over all partitions of
  \([a,b]\). Thus, \(V[cf;a,b]\leq |c|V\) so \(cf\) is b.v.\@ on \([a,b]\).

  For part (ii) given a partition \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of
  the interval \([a,b]\), by the triangle inequality we have
  \begin{align*}
    S_\Gamma
    &=\sum_{i=1}^n\left|(f(x_i)+g(x_i))-(f(x_{i-1})+g(x_{i-1}))\right|\\
    &=\sum_{i=1}^n\left|(f(x_i)-f(x_{i-1}))+(g(x_i)-g(x_{i-1}))\right|\\
    &\leq\sum_{i=1}^n|f(x_i)-f(x_{i-1})|+\sum_{i=1}^n|g(x_i)-g(x_{i-1})|\\
    &\leq V+V'.
  \end{align*}
  Thus, \(f+g\) is b.v.\@ on \([a,b\)]

  For part (iii) since \(f\) and \(g\) are b.v.\@ on \([a,b]\) by part (a)
  \(f\) and \(g\) are bounded on \([a,b]\) by, say, \(M\) and \(N\),
  respectively. Now, given a partition
  \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of \([a,b]\), by the triangle
  inequality we have
  \begin{align*}
    S_{\Gamma}
    &=\sum_{i=1}^n\left|f(x_i)g(x_i)-f(x_{i-1})g(x_{i-1})\right|\\
    &\begin{aligned} =\sum_{i=1}^n
      \left|f(x_i)g(x_i)\right.{}&{}\left.-f(x_{i-1})g(x_{i-1})\right.\\
      &\left.+f(x_i)g(x_{i-1})-f(x_i)g(x_{i-1})\right|
    \end{aligned}\\
    &\begin{aligned} =\sum_{i=1}^n
      \left|(f(x_i)g(x_i)\right.{}&{}\left.-f(x_i)g(x_{i-1}))\right.\\
      &\left.-(f(x_{i-1})g(x_{i-1})-f(x_i)g(x_{i-1}))\right|
    \end{aligned}\\
    &\begin{aligned}
      \leq\sum_{i=1}^n|f(x_i)g(x_i){}&{}-f(x_i)g(x_{i-1})|\\
      &+\sum_{i=1}^n|f(x_{i-1})g(x_{i-1})-f(x_i)g(x_{i-1})|
    \end{aligned}\\
    &=\sum_{i=1}^n|f(x_i)||g(x_i)-g(x_{i-1})|+\sum_{i=1}^n|g(x_{i-1})||f(x_i)-f(x_{i-1})|\\
    &=\sum_{i=1}^n
      M|g(x_i)-g(x_{i-1})|+\sum_{i=1}^n N|f(x_i)-f(x_{i-1})|\\
    &\leq MV'+NV.
  \end{align*}
  Thus, \(fg\) is b.v.\@ on \([a,b]\).

  Finally, for part (iv) suppose there exists \(\varepsilon>0\) such that
  \(|g(x)|\geq\varepsilon\) for all \(x\in[a,b]\). Then, given a partition
  \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of \([a,b]\), largely by the
  triangle inequality, we have
  \begin{align*}
    S_\Gamma
    &=\sum_{i=1}^n |f(x_i)/g(x_i)-f(x_{i-1})/g(x_{i-1})|\\
    &=\sum_{i=1}^n\left|\frac{f(x_i)g(x_{i-1})-
      f(x_{i-1})g(x_i)}{g(x_i)g(x_{i-1})}\right|\\
    &\leq\frac{1}{\varepsilon^2}\sum_{i=1}^n|f(x_i)g(x_{i-1})-f(x_{i-1})g(x_i)|\\
    &
      \begin{aligned}
        =\frac{1}{\varepsilon^2}\sum_{i=1}^n |f(x_i)g(x_{i-1}){}&{}-f(x_{i-1})g(x_{i-1})\\
        &-(f(x_{i-1})g(x_i)-f(x_{i-1})g(x_{i-1}))|
      \end{aligned}\\
    &\leq
      \frac{1}{\varepsilon^2}\sum_{i=1}^n|g(x_{i-1})||f(x_i)-f(x_{i-1})|
      +\frac{1}{\varepsilon^2}\sum_{i=1}^n|f(x_{i-1})||g(x_i)-g(x_{i-1})|\\
    &=\frac{1}{\varepsilon^2}\sum_{i=1}^nM_g|f(x_i)-f(x_{i})|
      +\frac{1}{\varepsilon^2}\sum_{i=1}^nM_f|g(x_i)-g(x_i)|\\
    &=\frac{1}{\varepsilon^2}M_g\sum_{i=1}^n|f(x_i)-f(x_{i})|
      +\frac{1}{\varepsilon^2}M_f\sum_{i=1}^n|g(x_i)-g(x_i)|\\
    &\leq\frac{1}{\varepsilon^2}(NV+MV')
  \end{align*}
  where, as above, \(f\) is bounded by \(M\) and \(g\) is bounded by
  \(N\). Thus, \(f/g\) is b.v.\@ on \([a,b]\).
  \\\\
  This concludes the proof of Theorem 2.1.
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 2, Ex.\@ 3]
  If \([a',b']\) is a subinterval of \([a,b]\) show that
  \(P[a',b']\leq P[a,b]\) and \(N[a',b']\leq N[a,b]\).
\end{problem}
\begin{solution}
  We will prove this by digging in to the definition of \(N\) and \(P\).
  Recall that given a partition \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) of
  the interval \([a,b]\), \(P\) and \(N\) are defined to be the supremum
  over the sum of the positive and, respectively, the sum negative terms of
  \(S_\Gamma\); that is, \(P\) and \(N\) are the supremum over every
  partition \(\Gamma\) of \([a,b]\) of
  \[
    P_\Gamma=\sum_{i=1}^n\left[f(x_i)-f(x_{i-1})\right]^+
    \quad\text{and}\quad
    N_\Gamma=\sum_{i=1}^n\left[f(x_i)-f(x_{i-1})\right]^-.
  \]

  Let \(f\colon[a,b]\to\bbR\) be a function of bounded variation on
  \([a,b]\) and let \([a',b']\) be a subinterval of \([a,b]\). Without loss
  of generality, we may assume that \([a',b']\) is strictly contained in
  \([a,b]\); that is, \(a'\neq a\) and \(b'\neq b\). We aim to show that
  \(P[a',b']\leq P[a,b]\) and \(N[a',b']\leq N[a,b]\). Since the argument
  for \(N\) is similar to that of \(P\), we will omit it here for the sake
  of brevity. Now, consider the closure of the complement of \([a',b']\) in
  \([a,b]\), \(\overline{[a,b]\setminus [a',b']}=[a,a']\cup[b',b]\). Since
  \([a,a']\), \([a',b']\) and \([b',b]\) are close intervals we may take
  partitions
  \begin{align*}
    \Gamma_a&=\{\,x_0<x_1\cdots<x_\ell\,\},\\
    \Gamma_{ab}&=\{\,x_\ell<x_{\ell+1}<\cdots<x_m\,\}
                 \shortintertext{and}
                 \Gamma_b&=\{\,x_m<x_{m+1}<\cdots<x_n\,\}
  \end{align*}
  of \([a,a']\), \([a',b']\) and \([b',b]\), respectively and extend this
  to a partition
  \[
    \Gamma=\{\,x_0<x_1<\cdots<x_\ell<x_{\ell+1}\cdots<x_m<x_{m+1}\cdots<x_n\,\}
  \]
  of \([a,b]\). Then, by the definition of \(N\) we have the string of
  inequalities
  \begin{align*}
    P_\Gamma&=\sum_{i=1}^n[f(x_i)-f(x_{i-1})]^+\\
            &\begin{aligned}
              =\sum_{i=1}^\ell&[f(x_i)-f(x_{i-1})]^+\\
              &+\sum_{i=\ell+1}^m[f(x_i)-f(x_{i-1})]^+\\
              &+\sum_{i=m+1}^n[f(x_i)-f(x_{i-1})]^+
            \end{aligned}\\
            &=P_{\Gamma_{ab}}+P_{\Gamma_a}+P_{\Gamma_b}\\
            &\leq P[a,b].
  \end{align*}
  Taking the supremum on the left, we have
  \[
    P[a,a']+P[a',b']+P[b',a']\leq P[a,b].
  \]
  Since \(P\) is strictly positive, it must be the case that
  \(P[a',b']\leq P[a,b]\).
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 2, Ex.\@ 11]
  Show that \(\int_a^bf\diff\varphi\) exists if and only if given
  \(\varepsilon>0\) there exists \(\delta>0\) such that
  \(\left|R_\Gamma-R_{\Gamma'}\right|<\varepsilon\) if
  \(|\Gamma|,|\Gamma'|<\delta\).
\end{problem}
\begin{solution}
  One direction is straightforward. Namely \(\impliedby\): suppose that
  given \(\varepsilon>0\) there exists \(\delta>0\) such that
  \(|R_\Gamma-R_{\Gamma'}|<\varepsilon\) whenever \(|\Gamma|\) and
  \(|\Gamma'|\) are less than \(\delta\). Let
  \({\{\Gamma_n\}}_{n=1}^\infty\) be a decreasing sequence of partitions
  (by which we mean \(\Gamma_n\subseteq\Gamma_{n+1}\) of \([a,b]\) such
  that \(|\Gamma_n|\to 0\). Then, by convergence, there exists \(N\in\bbN\)
  such that \(n\geq N\) implies \(|\Gamma_n|<\delta\). Then, for
  \(n,m\geq N\), we have
  \begin{align*}
    |R_{\Gamma_n}-R_{\Gamma_m}|<\varepsilon.
  \end{align*}
  Thus, by the Cauchy criterion for convergence, the sequence
  \({\{R_{\Gamma_n}\}}_{n=0}^\infty\) converges and its limit is by
  definition the Riemann--Stieltjes integral \(\int_a^b f\diff\varphi\).

  On the other hand \(\implies\): suppose that \(I=\int_a^b f\diff\varphi\)
  exists. Then given \(\varepsilon>0\) there exists \(\delta>0\) such that
  \(|I-R_\Gamma|<\varepsilon/2\) whenever \(|\Gamma|<\delta\). Let
  \(\Gamma\) and \(\Gamma'\) be two partitions of \([a,b]\) with norm
  \(|\Gamma|,|\Gamma'|<\delta\). Then we have
  \begin{align*}
    |R_\Gamma-R_{\Gamma'}|
    &=|R_\Gamma-I-(R_{\Gamma'}-I)|\\
    &\leq |R_\Gamma-I|+|R_{\Gamma'}-I|\\
    &<\frac{\varepsilon}{2}+\frac{\varepsilon}{2}\\
    &=\varepsilon.
  \end{align*}
  Thus, \(I\) satisfies the Cauchy condition.
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 2, Ex.\@ 13]
  Prove theorem (2.16).
\end{problem}
\begin{solution}
  Recall the statement of Theorem 2.16:
  \begin{quote}
    \begin{enumerate}[label=(\roman*)]
    \item If \(\int_a^b f\diff\varphi\) exists, then so do
      \(\int_a^bcf\diff\varphi\) and \(\int_a^b f\diff(c\varphi)\) for any
      constant \(c\), and
      \[
        \int_a^bcf\diff\varphi=\int_a^bf\diff(c\varphi)=c\int_a^bf\diff\varphi.
      \]
    \item If \(\int_a^b f_1\diff\varphi\) and \(\int_a^bf_2\diff\varphi\)
      both exist, so does \(\int_a^b\left(f_1+f_2\right)\diff\varphi\), and
      \[
        \int_a^b\left(f_1+f_2\right)\diff\varphi=\int_a^bf_1\diff\varphi+\int_a^bf_2\diff\varphi.
      \]
    \item If \(\int_a^bf\diff\varphi_1\) and \(\int_a^bf\diff\varphi_2\)
      both exist, so does
      \(\int_a^bf\diff\left(\varphi_1+\varphi_2\right)\), and
      \[
        \int_a^bf\diff\left(\varphi_1+\varphi_2\right)=\int_a^bf\diff\varphi_1+\int_a^bf\diff\varphi_2.
      \]
    \end{enumerate}
  \end{quote}
  \noindent%
  We prove this in (Roman) numerical order.

  For (i) suppose that \(I=\int_a^bf\diff\varphi\) exists. Then, given
  \(\varepsilon>0\), there exists \(\delta>0\) such that
  \(|I-R_\Gamma|<\varepsilon/|c|\) whenever \(\Gamma\) is a partition of
  \([a,b]\) with \(|\Gamma|<\delta\). We claim that
  \(\int_a^b cf\diff\varphi=|c|I\). Let
  \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\) be a partition \([a,b]\) with
  \(|\Gamma|<\delta\). Then the Riemann--Stieltjes sums \(R_\Gamma'\) of
  the pair \((cf,\varphi)\) associated to \(\Gamma\) give us the chain of
  inequalities
  \begin{align*}
    ||c|I-R_\Gamma'|
    &=\left| |c|I-
      \sum_{i=1}^n cf(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})]
      \right|\\
    &=|c|\left|
      \sum_{i=1}^n cf(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})]
      \right|\\
    &=|c||I-R_\Gamma|\\
    &<|c|\frac{\varepsilon}{|c|}\\
    &=\varepsilon.
  \end{align*}
  Thus, \(\int_a^b cf\diff\varphi\) is Riemann--Stieltjes integrable and
  its integral is equal to \(|c|I\). A similar argument shows that
  \(\int_a^b f\diff(c\varphi)\) is Riemann--Stieltjes integrable with
  integral \(|c|I\).

  For (ii) let \(I_1=\int_a^b f_1\diff\varphi\) and
  \(I_2=\int_a^b f_2\diff\varphi\). Then, we claim that
  \(I=\int_a^b (f_1+f_2)\diff\varphi\) exists and that \(I=I_1+I_2\). Since
  both \(I_1\) and \(I_2\) exist, given \(\varepsilon>0\) there exists
  \(\delta>0\) such that
  \[
    |I_1-R_\Gamma^1|<\frac{\varepsilon}{2}\qquad\text{and}\qquad
    |I_2-R_\Gamma^2|<\frac{\varepsilon}{2}
  \]
  whenever \(|\Gamma|<\delta\). Let \(\Gamma=\{\,x_0<x_1<\cdots<x_n\,\}\)
  be a partition of \([a,b]\) with \(|\Gamma|<\delta\). Then the
  Riemann--Stieltjes sums \(R_\Gamma\) of the pair \((f_1+f_2,\varphi)\)
  associated to \(\Gamma\) give is the following chain of inequalities
  \begin{align*}
    |(I_1+I_2)-R_{\Gamma}|
    &=\left|(I_1+I_2)
      -\sum_{i=1}^n (f_1(\xi_i)+f_2(\xi_i))[\varphi(x_i)-\varphi(x_{i-1})]
      \right|\\
    &=\left|
      I_1
      -\sum_{i=1}^n
      f_1(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})\right.\\
    &\qquad
      \left.+I_2
      -\sum_{i=1}^n f_2(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})]\right|\\
    &\leq
      \left|
      I_1
      -\sum_{i=1}^n
      f_1(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})
      \right|\\
    &\qquad
      +\left|
      I_2-\sum_{i=1}^n f_2(\xi_i)[\varphi(x_i)-\varphi(x_{i-1})]\right|\\
    &=|I_1-R_\Gamma^1|+|I_2-R_\Gamma^2|\\
    &<\frac{\varepsilon}{2}+\frac{\varepsilon}{2}\\
    &=\varepsilon.
  \end{align*}
  Thus, \(I\) exists and it is equal to the sum \(I_1+I_2\).

  Part (iii) is similar to part (ii) in the above equation except that
  instead of splitting the sum at \(f_1+f_2\) part, we split it at
  \(\varphi_1+\varphi_2\) part.
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
