\subsubsection{Homework 11}
\setcounter{exercise}{0}
\setcounter{equation}{0}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 7, Ex.\@ 11]
  Prove the following result concerning changes of variable. Let \(g(t)\)
  be monotone increasing and absolutely continuous on \([\alpha,\beta]\)
  and let \(f\) be integrable on \([a,b]\), \(a= g(\alpha)\),
  \(b= g(\beta)\). Then \(f(g(t))g'(t)\) is measurable and integrable on
  \([\alpha,\beta]\), and
  \[
    \int_a^b f(x)d x=\int_\alpha^\beta f(g(t))g'(t)\diff t.
  \]
  (Consider the case when \(f\) is the characteristic function of an
  interval, an open set, etc.)
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 7, Ex.\@ 15]
  Theorem 7.43 shows that a convex function is the indefinite integral of a
  monotone increasing function. Prove the converse: If
  \(\varphi(x)=\int_a^xf(t)\diff t+\varphi(a)\) in \((a,b)\) and \(f\) is
  monotone increasing, then \(\varphi\) is convex in \((a,b)\). (Use
  Exercise 14.)
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 5, Ex.\@ 8]
  Prove (5.49).
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 5, Ex.\@ 11]
  For which \(p\) does \(1/x\in L^p(0,1)\)? \(L^p(1,\infty)\)?
  \(L^p(0,\infty)\)?
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 5, Ex.\@ 12]
  Give an example of a bounded continuous \(f\) on \((0,\infty)\) such that
  \(\lim_{x\to\infty}f(x)=0\) but \(f\notin L^p(0,\infty)\) for any
  \(p>0\).
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 5, Ex.\@ 17]
  If \(f\geq 0\) and \(\omega(\alpha)\leq c(1+\alpha)^p\) for all
  \(\alpha>0\), show that \(f\in L^r\), \(0<r<p\).
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}[Wheeden \& Zygmund Ch.\@ 8, Thm.\@ 8.3]
  If \(f,g\in L^p(E)\), \(p>0\), then \(f+g\in L^p(E)\) and
  \(cf\in L^p(E)\) for any constant \(c\).
\end{problem}
\begin{solution}
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
