\subsubsection{Exam 1}
\setcounter{exercise}{0}
\setcounter{equation}{0}


\begin{problem}
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}
\hfill
\begin{enumerate}[label=(\roman*),noitemsep]
\item Show that if $B_r=\left\{\, x \in\bbR^n:| x |<r\,\right\}$,
  then there exists a constant $C$ such that $|B_r|=Cr^n$.
\\\\
(\emph{Hint}: Think of $B_r$ as $\left\{\,r x : x \in B_1\,\right\}$.)
\item Let $E\subset\bbR^n$ be a measurable set and let
  $\varphi_E\colon\bbR^n\to\bbR$ be defined
  $\varphi_E( x )=\left|E\cap B_{| x |}\right|$. Use part (i) to prove
    that $\varphi_E$ is continuous.
\end{enumerate}
\end{problem}
\begin{solution}
(i) To prove this result, we use the map constructed in Problem 1 of the
review sheet for Exam 1, the map $T\colon\bbR\times\bbR^n\to\bbR^n$. Set
$T_r\colon\bbR^n\to\bbR^n$ to be $T_r= T(r)$. Then, we claim
$B_r=T_r(B_1)$ and $|B_r|=|T_r(B_1)|$, which, as we saw in Problem 1 of the
review sheet, has measure $|B_1||r|^n$. Setting $C= |B_1|$, we have
$|B_r|=C|r|^n$ as desired.
\\\\
(ii) To prove that $\varphi_E$ is continuous, we provide an
$(\varepsilon,\delta)$-argument. Let $\varepsilon>0$ be given. We must show
that there exists $\delta>0$ such that $ y \in B( x ,\delta)$ implies
\begin{equation}
\label{eq:ex:1:1}
\left|\varphi_E( x )-\varphi_E( y )\right|<\varepsilon.
\end{equation}

First, note that since $ x \mapsto | x |$ is continuous and polynomials
$p\colon\bbR^n\to\bbR^n$ are continuous, then the composition
$ x \mapsto| x |^n$ is continuous. Therefore, there exists $\delta>0$
such that $ y \in B( x ,\delta)$ implies
\begin{equation}
\label{eq:ex:1:2}
\left|| x |^n-| y |^n\right|<\frac{\varepsilon}{C},
\end{equation}
where $C= |B_1|$.

Now, let $x\in\bbR^n$ and $ y \in B( x ,\delta)$ as above. Then, by
\eqref{eq:ex:1:2} we have
\begin{equation}
\label{eq:ex:1:3}
\begin{aligned}
\left|\varphi_E( x )-\varphi_E( y )\right|
&=\left|
\left|E\cap B_{| x |}\right|
-
\left|E\cap B_{| y |}\right|
\right|\\
&\leq\left| |B_{| x |}|-|B_{| y |}|\right|\\
&=C|\left| x |^n-| y |^n\right|\\
&\leq C\left[\frac{\varepsilon}{C}\right]\\
&=\varepsilon.
\end{aligned}
\end{equation}
It follows that $\varphi_E$ is continuous.
\end{solution}

\begin{problem}
Assume that $f\colon[a,b]\to\bbR$ is of bounded variation on $[a,b]$. Prove
that $f$ is measurable.
\end{problem}
\begin{solution}
By Jordan's theorem (Corollary 2.7 from \cite[Ch.\@ 2, p.\@
21]{wheeden-zygmund}), the function $f$ is of bounded variation on $[a,b]$
if and only if it can be written as the difference $f_1-f_2$ of two bounded
functions $f_1$ and $f_2$ that are monotone increasing on $[a,b]$. Then,
$f_1$ and $f_2$ are continuous a.e.\@ on $[a,b]$ and hence, are
measurable.
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
