\subsection{Homework 12}
\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 2]
Prove the converse of Hölder's inequality for $p=1$ and $\infty$. Show also
that for $1\leq p\leq\infty$, a real-valued measurable $f$ belongs to
$L^p(E)$ if $fg\in L^1(E)$ for every $g\in L^{p'}(E)$, $1/p+1/p'=1$. The
negation is also of interest: if $f\in L^p(E)$ then there exists $g\in
L^{p'}(E)$ such that $fg\notin L^1(E)$. (To verify the negation, construct
$g$ of the form $\sum a_kg_k$ satisfying $\int_E fg_k\to\infty$.)
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 3]
Prove Theorems 8.12 and 8.13. Show that Minkowski’s inequality for series
fails when $p<1$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 4]
Let $f$ and $g$ be real-valued and not identically $0$ (i.e., neither
function equals $0$ a.e.), and let $1<p<\infty$. Prove that equality holds
in the inequality $\left|\int fg\right|\leq\|f\|_p\|g\|_{p'}$ if and only if
$fg$ has constant sign a.e.\@ and $|f|^p$ is a multiple of $|g|^{p'}$ a.e.
\\\\
If $\|f+g\|_p=\|f\|_p+\|g\|_{p}$ and $g\neq 0$ in Minkowski's inequality,
show that $f$ is a multiple of $g$.
\\\\
Find analogues of these results for the spaces $\ell^p$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 5]
For $0<p\leq\infty$ and $0<|E|<\infty$, define
\[
N_p[f]\coloneqq\left(\frac{1}{E}\int_E|f|^p\right)^{1/p},
\]
where $N_\infty[f]$ means $\|f\|_\infty$. Prove that if $p_1<p_2$, then
$N_{p_1}[f]\leq N_{p_2}[f]$. Prove also that if $1\leq p\leq \infty$, then
$N_p[f+g]\leq N_p[f]+N_p[g]$, $(1/|E|)\int_E|fg|\leq N_p[f]N_{p'}[g]$,
$1/p+1/p'=1$, and $\lim_{p\to\infty} N_p[f]=\|f\|_\infty$. Thus, $N_p$
behaves like $\|\cdot\|_p$ but has the advantage of being monotone in
$p$. Recall Exercise 28 of Chapter 5.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 6]
\begin{enumerate}[label=(\alph*)]
\item Let $1\leq p_i$, $r\leq\infty$ and $\sum_{i=1}^k1/p_i=1/r$. Prove the
  following generalization of Hölder's inequality:
\[
\|f_1\dotsm f_k\|_r\leq\|f_1\|_{p_1}\dotsm\|f_k\|_{p_k}.
\]
\item Let $1\leq p<r<q\leq\infty$ and define $\theta\in(0,1)$ by
  $1/r=\theta/p+(1-\theta)/q$. Prove the interpolation estimate
\[
\|f\|_r\leq{\|f\|_p}^\theta{\|f\|_q}^{1-\theta}.
\]
In particular, if $A\coloneqq\max\left\{\|f\|_p,\|f\|_q\right\}$, then
$\|f\|_r\leq A$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 9]
If $f$ is real-valued and measurable on $E$, $|E|>0$, define its essential
infimum on $E$ by
\[
\essinf f\coloneqq\sup\left\{\,\alpha:\left|\{\,x\in
    E:f(x)<\alpha\,\}\right|=0\,\right\}.
\]
If $f\geq 0$, show that $\essinf_E f=(\esssup 1/f)^{-1}$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}[Wheeden \& Zygmund {\S}8, Ex.\@ 11]
If $f_k\to f$ in $L^p$, $1\leq p<\infty$, $g_k\to g$ pointwise, and
$\|g_k\|_\infty<M$ for all $k$, prove that $f_kg_k\to fg$ in $L^p$.
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
