\begin{problem}
  \begin{enumerate}[label=(\alph*),noitemsep]
  \item Show that for \(n=3\) the general solution to the wave equation
    \(u_{tt}-\Delta u=0\) with spherical symmetry about the origin has the
    form
    \[
      u=\tfrac{1}{r}F(r+t)+\tfrac{1}{r}G(r-t),\quad r=|x|,
    \]
    with suitable \(F\) and \(G\).
  \item Show that the solution with initial data of the form
    \[
      u(r,0)=0,\quad u_t(r,0)=h(r)
    \]
    (\(h\) is an even function of \(r\)) is given by
    \[
      u=\frac{1}{2r}\int_{r-t}^{r+t} \rho h(\rho)\diff \rho.
    \]
  \end{enumerate}
\end{problem}
\begin{solution}
  For part (a): We show that
  \[
    u=\tfrac{1}{r}F(r+t)+\tfrac{1}{r}G(r-t)
  \]
  is in fact a solution to the wave equation in
  \(\bbR^3\times(0,\infty)\). By direct calculation, we have
  \begin{align*}
    u_{tt}%
    &=\tfrac{1}{r}F''(r+t)+\tfrac{1}{r}G''(r-t),\\
    u_{r}
    &=\tfrac{1}{r}F'(r+t)+\tfrac{1}{r}G'(r-t)
      -\tfrac{1}{r^2}F(r+t)-\tfrac{1}{r^2}G(r-t),\\
    u_{rr}
    &=\tfrac{1}{r}F''(r+t)+\tfrac{1}{r}G''(r-t)\\
    &\phantom{{}={}\quad}-\tfrac{2}{r^2}F'(r+t)-\tfrac{2}{r^2}G'(r-t)\\
    &\phantom{{}={}\qquad}+\tfrac{2}{r^3}F(r+t)+\tfrac{2}{r^3}G(r-t),\\
    \Delta_{S^2}u&=0,
  \end{align*}
  and lastly,
  \begin{align*}
    \Delta u%
    &=\tfrac{\partial}{\partial r^2}u%
      +\tfrac{2}{r}\tfrac{\partial}{\partial r}u%
      +\tfrac{1}{r^2}\Delta_{S^2} u\\
    &=\tfrac{1}{r}F''(r+t)+\tfrac{1}{r}G''(r-t)\\
    &\phantom{{}={}\quad}-\tfrac{2}{r^2}F'(r+t)-\tfrac{2}{r^2}G'(r-t)
    +\tfrac{2}{r^3}F(r+t)+\tfrac{2}{r^3}G(r-t)\\
    &\phantom{{}={}\quad}
      +\tfrac{2}{r^2}F'(r+t)+\tfrac{2}{r^2}G'(r-t)
      -\tfrac{2}{r^3}F(r+t)-\tfrac{2}{r^3}G(r-t)\\
    &\phantom{{}={}}+0\\
    &=\tfrac{1}{r}F''(r+t)+\tfrac{1}{r}G''(r-t).
  \end{align*}
  Therefore, looking at the equation for \(u_{tt}\), we see that
  \[
    u_{tt}-\Delta u=0;
  \]
  i.e., \(u(r,t)\defeq\frac{1}{r}F(r+t)+\frac{1}{r}G(r-t)\) is a solution
  to the wave equation with \(F\) and \(G\) at least twice differentiable
  and such that they satisfy the initial conditions of the (nonhomogeneous)
  wave equation.

  We still need to show that if \(u\) is a solution to the wave equation
  with spherical symmetry it has the form prescribed above. We trust this
  can be done for now and return to this problem as time permits.
  \\\\
  For part (b): We show by a direct computation that
  \[
    u=\frac{1}{2r}\int_{r-t}^{r+t}\rho h(\rho)\diff\rho
  \]
  solve the initial-value problem for the wave equation. First, we compute
  the necessary partial derivatives of \(u\),
  \begin{align*}
    u_t(r,t)
    &=\tfrac{1}{2r}[(r+t)h(r+t)-(r-t)h(r-t)],\\
    u_{tt}(r,t)
    &=\tfrac{1}{2r}[(r+t)h'(r+t)+h(r+t)+(r-t)h'(r-t)+h(r-t)],\\
    u_r(r,t)
    &=\tfrac{1}{2r}[(r+t)h(r+t)-(r-t)h(r-t)]
      -\frac{1}{2r^2}\int_{r-t}^{r+t}\rho h(\rho)\diff\rho\\
    &=\tfrac{1}{2r}[(r+t)h(r+t)-(r-t)h(r-t)]
      -\tfrac{1}{r}u(r,t),\\
    u_{rr}(r,t)&=h'(r+t)-h'(r-t)-\tfrac{1}{r}u_r(r,t)+\tfrac{1}{r^2}u(r,t),
  \end{align*}
\end{solution}
\newpage

\begin{problem}
  Show that the solution \(w(x_1,t)\) of the initial-value problem for the
  \emph{Klein--Gordon equation}
  \begin{equation}
    \label{eq:9:klein-gordon-eq}
    \left\{
      \begin{aligned}
        w_{tt}=w_{x_1x_1}-\lambda^2w,\\
        w(x_1,0)=0,&&w_t(x_1,0)=h(x_1)
      \end{aligned}
    \right.
  \end{equation}
  is given by
  \[
    w(x_1,t)=\frac{1}{2}\int_{x_1-t}^{x_1+t}J_0(\lambda s) h(y_1)\diff y_1.
  \]
  Here \(s^2=t^2-(x_1-y_1)^2\), while \(J_0\) denotes the Bessel function
  defined by
  \[
    J_0(z)\defeq \frac{2}{\pi}\int_0^{\frac{\pi}{2}}\cos(z\sin\theta)\diff\theta.
  \]

  \noindent (\emph{Hint:} Descend to \eqref{eq:9:klein-gordon-eq} from the
  two-dimensional wave equation satisfied by
  \[
    u(x_1,x_2,t)=\cos(\lambda x_2)w(x_1,t).\text{)}
  \]
\end{problem}
\begin{solution}
  Taking the hint, the equation \(u(x_1,x_2,t)\defeq\cos(\lambda
  x_2)w(x_1,t)\) satisfies the two-dimensional wave equation as we will
  shortly see. First, let us compute find the necessary partial derivatives
  of \(u\),
  \begin{align*}
    u_{tt}&=\cos(\lambda x_2)w_{tt}(x_1,t),\\
    u_{x_1}&=\cos(\lambda x_2)w_{x_1}(x_1,t),\\
    u_{x_1x_1}&=\cos(\lambda x_2)w_{x_1x_1}(x_1,t),\\
    u_{x_2}&=-\lambda\sin(\lambda x_2)w(x_1,t),\\
    u_{x_2x_2}&=-\lambda^2\cos(\lambda x_2)w(x_1,t).
  \end{align*}
  Then, by \eqref{eq:9:klein-gordon-eq} together with the equations above
  we have
  \[
    \left\{
      \begin{aligned}
        u_{tt}-\Delta u=\cos(\lambda x_2)(w_{tt}-w_{x_1x_1}+\lambda^2 w)=0
        &&&\text{in \(\bbR^3\times(0,\infty)\),}\\
        u=0,\quad u_t=\cos(\lambda x_2)h(x_1)&&&\text{on
          \(\bbR^3\times\{\,t=0\,\}\).}
      \end{aligned}
    \right.
  \]
  By Kirchhoff's formula,
  \[
    u(x_1,x_2,t)=\fint_{\partial
      B(x_1,x_2,t)}t\cos(\lambda y_2)h(y_1)\diff S(y_1,y_2)
  \]
  solves the above initial-value problem. We can rewrite the above surface
  integral as the double integral
  \begin{align*}
    u(x_1,x_2,t)
    &=\frac{1}{4\pi}\int_{\partial B(0,0,1)}
      \cos(\lambda (x_2+tz_2))h(x_1+tz_1)\diff S(z)\\
    &=
  \end{align*}
\end{solution}
\newpage

\begin{problem}
  Let \(u\) solve
  \[
    \left\{
      \begin{aligned}
        u_{tt}-\Delta u=0\phantom{,}&&&&&\text{in \(\bbR^3\times(0,\infty)\),}\\
        u=g,&&u_t=h&&&\text{on \(\bbR^3\times\{\,t=0\,\}\)}
      \end{aligned}
    \right.
  \]
  where \(g\) and \(h\) are smooth and have compact support. Show there
  exists a constant \(C\) such that
  \[
    |u(x,t)|\leq \frac{C}{t}\quad(x\in\bbR^3,\,t>0).
  \]
\end{problem}
\begin{solution}
  Since \(h\) and \(g\) are compactly supported, we can find sufficiently
  large balls about their supports \(B(x,R)\) independent of the size of
  \(B(x,t)\). Now, by Kirchhoff's formula we have
  \begin{align*}
    |u(x,t)|
    &=\left|
      \fint_{\partial B(x,t)} th(y)+g(y)+Dg(y)\cdot(y-x)\diff S(y)
      \right|\\
    &=\frac{1}{4\pi t^2}\left|
      \int_{\partial B(x,t)} th(y)+g(y)+Dg(y)\cdot(y-x)\diff S(y)
      \right|\\
    &\leq\frac{1}{4\pi t^2}\int_{\partial B(x,R)}t\sup\{|h|,|g|,
      |Dg|\}\diff S(y)\\
    &=\frac{M|\partial B(x,R)|}{4\pi t}\\
    &=\frac{C}{t},
  \end{align*}
  where \(M<\infty\) since \(h\) and \(g\) are smooth and compactly
  supported (and hence \(h\), \(g\), and \(|Dg|\) achieve their supremum in
  \(B(x,R)\)).
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA523-HW-Current"
%%% End:
