\chapter{Covering Space Problems}
Compiled from Prof.\,McClure's old quals.
\begin{problem}

\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}
\section{Kyle's Stuff}
\begin{problem}[No.\,5]
Let $X$ be a topological space and let $x_0\in X$. Let $U$ and $V$ be open
sets containing $x_0$, and suppose that the hypotheses of the seifert--van
Kampen theorem are satisfied. Let $i_1\colon U\cap V\to U$, $i_2\colon
U\cap V\to V$, $j_1\colon U\to X$, and $j_2\colon V\to X$ be the inclusion
maps. Suppose that $\left(i_1\right)_*\colon \pi_1(U\cap
V,x_0)\to\pi_1(U,x_0)$ is onto. Prove, using the Seifert--van Kampen
theorem, that $\left( j_2 \right)_*\colon\pi_1(V,x_0)\to\pi_1(X,x_0)$ is
onto.
\end{problem}
\begin{proof}
We use the classical Seifert--van Kampen theorem (Theorem 70.2). Suppose
$\left(i_1\right)_*\colon\pi_1(U\cap V,x_0)\to\pi_1(U,x_0)$ is onto. Then
for every element $\gamma\in\pi_1(U,x_0)$,
$\gamma=\left(i_1\right)_*(\gamma')$ for some element
$\gamma'\in\pi_1(U\cap V,x_0)$. Now, let $\gamma''\in\pi_1(X,x_0)$. By the
classical Seifert--van Kapmen theorem, the map
\[
j\colon\pi_1(U,x_0)*\pi_1(V,x_0)\longrightarrow\pi_1(X,x_0)
\]
is surjective and its kernel is the leans normal subgroup $N$ of the free
product that contanis all elements represented by words of the form
$(i_1(g)^{-1},i_2(g))$.
\end{proof}
\begin{problem}[No.\,6]
As in 5., but instead suppose that $\left( i_1 \right)_*\colon\pi_1(U\cap
V,x_0)\to\pi_1(X,x_0)$ is an isomorphism. Prove, using the Seifert--van
Kampen theorem, that there is a homomorphism
$\Phi\colon\pi_1(X,x_0)\to\pi_1(V,x_0)$ for which
$\Phi\circ\left(j_2\right)_*$ is the identity map of $\pi_1(V,x_0)$.
\end{problem}
\begin{proof}

\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA571-Quals"
%%% End:
