\chapter{August 2014}
\begin{problem}
Let $X$ be a topological space, let $A$ be a subset of $X$, and let $U$ be
an open subset of $X$. Prove that $U\cap \bar A\subset\overline{U\cap A}$.
\end{problem}
\begin{proof}
Let $x\in U\cap\bar A$. Then $x\in U$ and $x\in\bar A$. This means that,
since $U$ is open, by Lemma C there exist an open neighborhood $V$ of $x$
such that $V\subset U$. Moreover, since $x\in\bar A$, $V'\cap
A\neq\emptyset$ for every open neighborhood $V'$ of $x$. In particular,
$V\cap A\neq\emptyset$. Thus, we have $V\cap U\neq\emptyset$ and $V\cap
A\neq\emptyset$ so $V\cap(U\cap A)\neq\emptyset$.
\end{proof}

\begin{problem}
Let $X$ be the following subspace of $\bfR^2$:
\[
((0,1]\times[0,1])\cup([2,3)\times[0,1]).
\]
Let $\sim$ be the equivalence relation on $X$ with $(1,t)\sim(2,t)$ (that
is $(s,t)\sim(s',t')\iff(s,t)=(s',t')$ or $t=t'$ and $\{s,s'\}=\{1,2\}$;
you do \emph{not} have to prove that this is an equivalence
relation). Prove that $X/{\sim}$ is homeomorphic to
$(0,2)\times[0,1]$. (\emph{Hint}: construct maps in both directions).
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that there is an equivalence relation $\sim$ on the interval $[0,1]$
such that $[0,1]/{\sim}$ is homeomorphic to $[0,1]\times[0,1]$. As part of
your proof \emph{explain} how you are using one or more properties of the
quotient topology.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $D$ be the closed unit disk in $\bfR^2$, that is, the set
\[
\left\{\,(x,y)\;\middle|\;x^2+y^2\leq 1\,\right\}.
\]
Let $E$ be the open unit disk
\[
\left\{\,(x,y)\;\middle|\;x^2+y^2<1\,\right\}.
\]
Let $X$ be the one-point compactification of $E$, and let $f\colon D\to X$
be the map defined by
\[
f(x,y)=
\begin{cases}
(x,y)&\text{if $x^2+y^2<1$}\\
\infty&\text{if $x^2+y^2=1$.}
\end{cases}
\]
Prove that $f$ is continuous.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $X$ and $Y$ be homotopy-equivalent topological spaces. Suppose that $X$
is path-connected. Prove that $Y$ is path-connected.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $a$ and $b$ denote the points $(-1,0)$ and $(1,0)$ in $\bfR^2$. Let
$x_0$ denote the origin $(0,0)$. Use the Seifert--van Kampen theorem to
calculate $\pi_1\left(\bfR^2-\{a,b\},x_0\right)$. You may not use any other
method.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $p\colon E\to B$ be a covering map with $B$ locally connected, and let
$x\in B$. Prove that $x$ has a neighborhood $W$ with the following
property: for every connected component $C$ of $p^{-1}(W)$, the map
$p\colon C\to W$ is a homeomorphism.
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA571-Quals"
%%% End:
