\chapter{Homework}
Here is this week's WebAssign homework and outlines of the solutions to the
problems.

I have three sections the first one beginning at \textbf{1:30 PM} and the
last one at 3:30 pm in \textbf{REC 315}. Students should only ask questions
about homework \# 25, 26, and 27.

\bigskip

\textkr{성준아}, here is the material you need to know to give this
recitation tomorrow. The subject for tomorrow's recitation are the
different convergence tests for series found in sections \S 11.6
to 11.8 of Stewart's book. Here is a rough summary of the material on those
sections:
\section*{Stewart \S 11.6--\S 11.8 Summary}
\subsection*{\S 11.6: Absolute Convergence and the Ratio and Root Tests}
\begin{definition}
A series $\sum a_n$ is called \emph{absolutely convergent} if the series of
absolute values $\sum |a_n|$ is convergent.
\end{definition}
\begin{definition}
A series $\sum a_n$ is called \emph{conditionally convergent} if it is
convergent but not absolutely convergent.
\end{definition}
\begin{theorem}
If a series $\sum a_n$ is absolutely convergent, then it is conditionally
convergent.
\end{theorem}
\begin{theorem}[The Ratio Test]
\begin{enumerate}[label=\textnormal{(\roman*)},noitemsep]
\item If $\lim\left|a_{n+1}/a_n\right|=L<1$, then the series
  $\sum a_n$ is absolutely convergent (and therefore convergent).
\item If $\lim\left|a_{n+1}/a_n\right|=L>1$ or $\infty$, then the series
  $\sum a_n$ is divergent.
\item If $\lim\left|a_{n+1}/a_n\right|=L=1$, the \emph{ratio test} is
  inconclusive; that is, conclusion can be drawn about the convergence or
  divergence of $sum a_n$.
\end{enumerate}
\end{theorem}
\begin{theorem}[The Root Test]
\begin{enumerate}[label=\textnormal{(\roman*)},noitemsep]
\item If $\lim\sqrt[n]{|a_n|}=L<1$, the series $\sum a_n$ is is absolutely
  convergent (and therefore convergent).
\item If $\lim\sqrt[n]{|a_n|}=L>1$ or $\infty$, then the series $\sum
  a_n$ is divergent.
\item If $\lim\sqrt[n]{|a_n|}=L=1$, the \emph{root test} is inconclusive.
\end{enumerate}
\end{theorem}
\subsection*{\S 11.7: Strategy for Testing Series}
\begin{enumerate}[label=\textbf{\arabic*.}]
\item If the series is of the form $\sum 1/n^p$, it is a $p$-series, which
  we know to be convergent if $p>1$ and divergent if $p\leq 1$.
\item If the series has the form $\sum ar^{n-1}$ or $\sum ar^n$, it is a
  geometric series, which converges if $|r|<1$ and diverges if $|r|\geq
  1$.
\item If the series has a form similar to a $p$-series or geometric series,
  then one of the comparison tests should be considered.
\item If you can see $\lim a_n\neq 0$, then the test for divergence should
  be used.
\item If the series has the form $\sum(-1)^{n+1} b_n$ or $\sum (-1)^n b_n$,
  the alternating test is an obvious choice.
\item Series that involve factorials and other products are handled
  conveniently with the ratio test.
\item If $a_n$ has the form ${b_n}^n$, then the root test may be useful.
\item If $a_n=f(n)$, where $\int_1^\infty f(x)\diff x$ is easily evaluated,
  the integral test is effective.
\end{enumerate}
\subsection*{\S 11.8: Power Series}
\begin{definition}
A \emph{power series} is a series of the form
\[
\sum_{n=0}^\infty c_nx^n=c_0+c_1x+c_2x^2+\dotsb
\]
where $x$ is a variable and the $c_n$'s are constants called
\emph{coefficients} of the series.
\end{definition}
\begin{definition}
More generally, a series of the form
\[
\sum_{n=0}^\infty c_n(x-a)^n=c_0+c_1(x-a)+c_2+(x-a)^2+\dotsb
\]
is called a \emph{power series in $(x-a)$} or a \emph{power series centered
at $a$} or a \emph{power series series about $a$}.
\end{definition}
\begin{theorem}
For a given power series $\sum c_n(x-a)^n$ there are only three
possibilities:
\begin{enumerate}[label=\textnormal{(\roman*)},noitemsep]
\item The series converges only when $x=a$.
\item The series converges for all $x$.
\item There is a positive number $R$ such that the series converges if
  $|x-a|<R$ and diverges if $|x-a|>R$.
\end{enumerate}
\end{theorem}
The number $R$ in (iii) is called the \emph{radius of convergence} of the
power series. By convention, $R=0$ in case (i) and $\infty$ in case (ii).
The \emph{interval of convergence} of a power series is the interval that
consists of all values of $x$ for which the series converges.
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
&Series&Radius of convergence&Interval of convergence\\
\hline
&&&\\
Geometric series&$\displaystyle\sum x^n$&$R=1$&$(-1,1)$\\&&&\\
Example 1&$\displaystyle\sum n!x^n$&$r=0$&$\{0\}$\\&&&\\
Example 2&$\displaystyle\sum\frac{(x-3)^n}{n}$&$R=1$&$[2,4)$\\&&&\\
Example
  3&$\displaystyle\sum\frac{(-1)^nx^{2n}}{2^{2n}(n!)^2}$&$R=\infty$&$(-\infty,\infty)$\\
&&&\\
\hline
\end{tabular}
\end{center}
Now
\newpage
\section*{Homework 25}
\begin{problem}[WebAssign HW 25, \# 1]
Determine whether the series is absolutely convergent, conditionally
convergent, or divergent.
\[
\sum_{n=1}^\infty\frac{n}{\sqrt{n^3-9}}.
\]
\end{problem}
\begin{problem}[WebAssign HW 25, \# 2]
Determine whether the series is absolutely convergent, conditionally
convergent, or divergent.
\[
\sum_{n=1}^\infty\frac{\sin 7n}{5^n}.
\]
\end{problem}
\begin{problem}[WebAssign HW 25, \# 3]
Determine whether the series is absolutely convergent, conditionally
convergent, or divergent.
\[
\sum_{n=1}^\infty\frac{12^n}{(n+1) 5^{2n+1}}.
\]
\end{problem}
\begin{problem}[WebAssign HW 25, \# 4]
Determine whether the series is absolutely convergent, conditionally
convergent, or divergent.
\[
\sum_{n=2}^\infty\frac{(-1)^n}{\ln 6n}.
\]
\end{problem}
\begin{problem}[WebAssign HW 25, \# 5]
The terms of a series are defined recursively by the equations
\[
a_1=4\qquad a_{n+1}=\frac{7n+1}{3n+9}\cdot a_n.
\]
Determine whether $\sum a_n$ is absolutely convergent, conditionally
convergent, or divergent.
\end{problem}

\section*{Homework 26}
\begin{problem}[WebAssign HW 26, \# 1]
Determine whether the series is absolutely convergent, conditionally
convergent, or divergent.
\[
\sum_{n=1}^\infty\left(\frac{n^2+4}{5n^2+2}\right)^n.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 2]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty\frac{1}{n+6^n}
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 3]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty\frac{(4n+1)^n}{n^{5n}}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 4]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty (-1)^n\frac{5n}{n+3}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 5]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty\frac{n^28^{n-1}}{(-9)^n}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 6]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty \frac{1}{8n+5}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 7]
Test the series for convergence or divergence.
\[
\sum_{n=2}^\infty\frac{1}{n\sqrt{\ln 7n}}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 8]
Test the series for convergence or divergence.
\[
\sum_{k=1}^\infty\frac{6^kk!}{(k+2)!}.
\]
\end{problem}
\begin{problem}[WebAssign HW 26, \# 9]
Test the series for convergence or divergence.
\[
\sum_{n=1}^\infty\frac{6n!}{e^{n^2}}.
\]
\end{problem}
\section*{Homework 27}
\begin{problem}[WebAssign HW 27, \# 1]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=2}^\infty\frac{x^{n+1}}{2n!}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 2]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty\frac{8^nx^n}{n^3}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 3]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty\frac{(-5)^n}{n\sqrt{n}}x^n.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 4]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=2}^\infty(-1)^n\frac{x^n}{12^n\ln n}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 5]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=0}^\infty\frac{(x-4)^n}{n^6+1}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 6]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty\frac{6^n(x+4)^n}{\sqrt{n}}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 7]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty\frac{(x-6)^n}{n^n}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 8]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty\frac{n}{b^n}(x-a)^n, \qquad b>0.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 9]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=1}^\infty n!(3x-1)^n.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\begin{problem}[WebAssign HW 27, \# 10]
Find the radius of convergence, $R$, of the series.
\[
\sum_{n=2}^\infty\frac{x^{6n}}{n(\ln n)^8}.
\]
Find the interval, $I$, of convergence of the series.
\end{problem}
\newpage
\chapter{Relevant Exam Problems}
If you run out of things to talk about within the first fifteen to twenty
minutes, talk about these
\begin{problem}[Exam 3, Spring 2015, \# 2]
The series
\[
\sum_{n=1}^\infty\frac{1}{\left( n^{3\alpha}+9 \right)^{1/8}}
\]
if and only if $\alpha$ is?
\end{problem}
\begin{proof}[Solution]
Note that
\[
(n^{3\alpha}+9)^{1/8}>n^{3\alpha/8}
\]
so that
\[
\frac{1}{(n^{3\alpha}+9)^{1/8}}<\frac{1}{n^{3\alpha/8}}
\]
the left of which forms a $p$-series with $p=3\alpha/8$. For this
$p$-series to converge $3\alpha/8>1$ so $\alpha>8/3$. This narrows down the
answer to exactly one of the choices, in the exam namely C.
\end{proof}
\begin{problem}[Exam 3, Spring 2015, \# 3]
Test the following series for convergence or divergence.
\begin{enumerate}[label=(\alph*)]
\item $\displaystyle\sum_{n=1}^\infty n\sin\left(\frac{1}{n}\right)$.
\item $\displaystyle\sum_{n=1}^\infty(-1)^n\arctan(\pi/2n)$.
\item $\displaystyle\sum_{n=1}^\infty\frac{n^2+9}{(n^3+4)\sqrt{n}}$.
\end{enumerate}
\end{problem}
\begin{proof}[Solution]
In the case of (a), you've seen time and time again that the limit of the
sequence $a_n\coloneqq n\sin(1/n)$ goes to $1$ as $n\to\infty$ (if you are
not convinced, use l'Hôpital's rule on the limit $x\sin(1/x)$ as
$x\to\infty$). In particular, since the limit is nonzero, this series
cannot converge.

In the case of (b), note that $a_n\coloneqq \arctan(\pi/2 n)$ fails to
satisfy the second condition for the alternating series test, namely, since
$\pi/2n\to \infty$ as $n\to\infty$, we have $\arctan(\pi/2 n)\to\pi/2\neq
0$, so the series must diverge.

Lastly, in the case of (c), we have
\[
(n^3+4)\sqrt{n}>n^3\sqrt{n}=n^{7/2}
\]
so
\begin{align*}
\sum_{n=1}^\infty\frac{n^2+9}{(n^3+4)\sqrt{n}}
&<\sum_{n=1}^\infty\frac{n^2+9}{n^{7/2}}\\
&=\sum_{n=1}^\infty\frac{n^2}{n^{7/2}}+
9\sum_{n=1}^\infty\frac{1}{n^{7/2}}\\
&=\underbrace{\sum_{n=1}^\infty\frac{1}{n^{5/2}}}_{S_1}+
9\underbrace{\sum_{n=1}^\infty\frac{1}{n^{7/2}}}_{S_2}
\end{align*}
Both of which are $p$-series with $p_1=5/2>1$ and $p_2=7/2>1$,
respectively. This implies that $S_1$ and $S_2$ converge so by the (as
Stewart calls it) direct comparison test, the original series
converges. Therefore, the letter answer to this problem is D.

\textbf{Correction!!!} It seems the answer is actually C and the difference
here is that the test writer is claiming (b) converges. The only
explanation can see for this is that he meant to write
$\arctan(\pi/(2n))$. In this case the conditions of the alternating series
tests are satisfied since (i) $\frac{\diff}{\diff
  x}(\arctan(\pi/(2x)))=-\pi x^{-2}/(2+\pi^2 x^{-2}/2)<0$ on $1\leq
x\leq\infty$ so the terms of the series are decreasing and (ii) $\lim
\arctan(\pi/(2n))=0$ since $\pi/(2n)\to 0$ as $n\to 0$ and $\arctan(0)=0$.
\end{proof}
\begin{problem}[Exam 3, Spring 2015, \# 7]
Suppose that the powers
\[
\sum_{n=0}^\infty c_n(x-5)^n
\]
converges when $x=2$ and diverges when $x=10$.

From the above information, which of the following statements can we
conclude to be true?
\begin{enumerate}[label=\MakeUppercase{\roman*}.]
\item The radius of convergence $R$ satisfies $3\leq R\leq 5$.
\item We \emph{cannot} determine the interval of convergence from the above
  information only.
\item The derivative of the power series is $\sum_{n=1}^\infty
  nc_n(x-5)^{n-1}$ which converges when $x=3$.
\end{enumerate}
\end{problem}
\begin{proof}[Solution]
For case I just recall what it means to be the radius of
convergence. It's a real number $R\geq 0$ such that $|x-5|<R$ implies that
the series converges. In this case, we are told that for $x=2$ and for
$x=10$, which tells us that the series converges for $|x-5|\leq 3$ and
diverges $|x-5|\geq 5$ and so we know that the radius of convergence cannot
exceed $5$ and is at least as big as $3$. Hence, I holds.

For case II, the information cannot be used to determine the interval of
convergence since, given that $3\leq R\leq 5$ is true, the interval of
convergence may be any one of the intervals with $(2,8)$ or $(0,10)$ as
endpoints (possibly including the endpoints).

For case III, given that we know $|2-5|\leq 3$ implies $\sum c_n(2-5)^n$
converges, by the limit comparison test (as Stewart calls it), since
$c_n/(nc_n)\to 0$  as $n\to\infty$, this implies that $\sum nc_n(2-5)^n$
converges. Therefore, the letter answer is D. (I and II are true).
\end{proof}

\begin{center}
{\large\textkr{고마워,성준!}}
\end{center}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA166-Recitation"
%%% End:
