% #7 - Due Feb. 29 Read Sections 5.1-2. Chapter 4: #9, 11, 15, 18 (ignore
% the second part concerning convergence in measure);
\begin{problem}[Wheeden \& Zygmund {\S}4, Ex.\@ 9]
\begin{enumerate}[label=(\alph*)]
\item Show that the limit of a decreasing (increasing) sequence of
  functions usc (lsc) at $\bfx_0$ is usc (lsc) at $\bfx_0$. In particular,
  the limit of a decreasing (increasing) sequence of functions continuous
  at $\bfx_0$ is usc (lsc) at $\bfx_0$.
\item Let $f$ be usc and less than $\infty$ on $[a,b]$. Show that there
  exists continuous $f_k$ on $[a,b]$ such that $f_k\searrow f$.
\end{enumerate}
\end{problem}
\begin{proof}
(a) Without loss of generality, assume that the sequence of $f_k$'s are
define in all of $\bfR^n$. Suppose
$\left\{f_k\colon\bfR^n\to\bfR\right\}_{k=1}^\infty$ is a sequence of
decreasing functions which are usc at $\bfx_0$ and put $f\coloneqq\lim
f_k$. Then, since the $f_k$'s are decreasing, we have
\begin{equation}
  \label{eq:decreasing-estimate-on-f}
f(\bfx)\leq f_k(\bfx)\qquad\text{for all $\bfx\in\bfR^n$.}
\end{equation}
Moreover, $f_k$ being usc at $\bfx_0$ means that for any sequence
$\bfx\to\bfx_0$, $\varlimsup_{\bfx\to\bfx_0} f_k(\bfx)\leq
f_k(\bfx_0)$. Thus, by \eqref{eq:decreasing-estimate-on-f} we have
\begin{equation}
  \label{eq:limit-estimate-on-f}
\varlimsup_{\bfx\to\bfx_0}f(\bfx)\leq\varlimsup_{\bfx\to\bfx_0}
f_k(\bfx)\leq f_k(\bfx_0).
\end{equation}
Now, let $k\to\infty$ and we have $\varlimsup_{\bfx\to\bfx_0} f(\bfx)\leq
f(\bfx_0)$. Thus, $f$ is usc at $\bfx_0$.
\\\\
(b) Suppose $f$ is usc and bounded on $[a,b]$.
\end{proof}
\newpage

\begin{problem}[Wheeden \& Zygmund {\S}4, Ex.\@ 11]
Let $f$ be defined on $\bfR^n$ and let $B(\bfx)$ denote the open ball
$\left\{\,\bfy\;\middle|\;\left|\bfx-\bfy\right|<r\,\right\}$ with center
$\bfx$ and fixed radius $r$. Show that the function
$g(\bfx)=\sup\left\{\,f(\bfy)\;\middle|\;\bfy\in B(\bfx)\,\right\}$ is lsc
and the function $h(\bfx)=\inf\left\{\,f(\bfy)\;\middle|\;\bfy\in
B(\bfx)\,\right\}$ is usc on $\bfR^n$. Is the same true for the closed ball
$\left\{\,\bfy\;\middle|\;\left|\bfx-\bfy\right|\leq r\,\right\}$?
\end{problem}
\begin{proof}
Fix $\bfx_0\in\bfR^n$. Given any sequence $\bfx\to\bfx_0$, we must show
that $\varliminf_{x\to\bfx_0} g(\bfx)\geq g(\bfx_0)$. Take any such
sequence $\bfx_k\to\bfx_0$ then convergence of such sequence means that for
any $\varepsilon>0$, the sequence $\{\bfx_k\}$ is eventually in
$B_\delta(\bfx_0)$, i.e., for some index $N\in\bfN$ depending on
$\varepsilon$, $\bfx_n\in B_\delta(\bfx_0)$ for all $n\geq N$. Hence, we
have $g(\bfx_0)\leq f(\bfx_n)\leq g(\bfx_n)$. Passing to the $\limsup$, we
have $g(\bfx_0)\leq\varlimsup_{\bfx_k\to\bfx_0} g(\bfx)$. Thus, $g$ is lsc.

Fix $\bfx_0\in\bfR^n$. Then, given any sequence $\bfx_k\to\bfx_0$ we must
show that $\varliminf_{\bfx_k\to\bfx_0}g(\bfx_k)\geq g(\bfx_0)$. Let
$\bfx_k\to\bfx_0$ be a convergent sequence then convergence of said
sequence means that for any $\varepsilon>0$, $\bfx_k$ is eventually in
$B_\varepsilon(\bfx_0)$, i.e., for some index $N\in\bfN$ depending on
$\varepsilon$, $\bfx_n\in B_\varepsilon(\bfx_0)$ for $n\geq N$. By the
definition of the supremum we have there exists $\bfy\in B(\bfx_0)$ such
that $g(\bfx_0)<f(\bfy)+\varepsilon$.
\end{proof}
\newpage

\begin{problem}[Wheeden \& Zygmund {\S}4, Ex.\@ 15]
Let $\left\{f_k\right\}$ be a sequence of measurable functions defined on a
measurable set $E$ with $|E|<\infty$. If $|f_k(M)|\leq M<\infty$ for all
$k$ for each $\bfx\in E$, show that given $\varepsilon>0$, there is closed
$F\subset E$ and finite $M$ such that $|E\minus F|<\varepsilon$ and
$|f_k(\bfx)|\leq M$ for all $\bfx\in F$.
\end{problem}
\begin{proof}
Define $f\coloneqq\sup|f_k|$. Note that, since $|f_k|=f^++f^-$ and $f^+$
and $f^-$ are measurable, $|f_k|$ is measurable hence, by 4.11, $f$ is
measurable. Now, given $\varepsilon>0$ by Lusin's theorem $f$ has the
$\calC$-property on $E$, i.e., there exists a closed subset $F$ of $E$ such
that $|E\minus F|<\varepsilon/2$ and $f$ is continuous when restricted to
$F$. Take the $\delta>0$ such that
$\bigl|E\minus\overline{B_\delta(\mathbf{0})}\bigr|<\varepsilon/2$. Then
$F\cap\overline{B_\delta(\mathbf{0})}$ is closed and compact and we have
\begin{align*}
\left|E\minus(F\cap\overline{B_\delta(\mathbf{0})})\right|
&=\left|E\minus F\cup
  \bigl(E\minus\overline{B_\delta(\mathbf{0})}\bigr)\right|\\
&\leq|E\minus F|+\bigl|E\minus\overline{B_\delta(\mathbf{0})}\bigr|\\
&<\varepsilon.
\end{align*}
By Problem 6.2 (W\&Z, 4.7) $f$ achieves its maximum $M$ on
$F\cap\overline{B_\delta(\mathbf{0})}$. Thus, $|f_k|\leq M$ for all
$\bfx\in F\cap\overline{B_\delta(\mathbf{0})}$.
\end{proof}
\newpage

\begin{problem}[Wheeden \& Zygmund {\S}4, Ex.\@ 18]
If $f$ is measurable on $E$, define
$\omega_f(a)=\left|\left\{\,f>a\,\right\}\right|$ for
$-\infty<a<\infty$. If $f_k\nearrow f$, show that
$\omega_{f_k}\nearrow\omega_f$. If $f_k\to f$, show that
$\omega_{f_k}\to\omega_f$ at each point of continuity of $\omega_f$. [For
the second part, show that if $f_k\to f$, then
$\varlimsup_{k\to\infty}\omega_{f_k}(a)\leq\omega_f(a-\varepsilon)$ and
$\varliminf_{k\to\infty}\omega_{f_k}(a)\geq\omega_f(a+\varepsilon)$ for every
$\varepsilon>0$.]
\end{problem}
\begin{proof}
\end{proof}
\newpage

% (Chapter 5: # 1, 2, 3, 4).
\begin{problem}[Wheeden \& Zygmund {\S}5, Ex.\@ 1]
If $f$ is a simple measurable function (not necessarily positive) taking
values $a_j$ on $E_j$, $j=1,...,N$, show that $\int_E f=\sum_{j=1}^N
a_j\left|E_j\right|$. [Use (5.24)].
\end{problem}
\begin{proof}
Since $f$ is a simple measurable function $E_k\cap E_\ell=\emptyset$ for
$k\neq\ell$. Since $E\coloneqq\bigcup_{j=1}^N E_j$ is countable, by 5.24 we
have
\[
\int_E f=\sum_{j=1}^N\int_{E_j} f=\sum_{j=1}^N\int a_j\chi_{E_j}=\sum_{j=1}^N a_j|E_j|.
\]
\end{proof}
\newpage

\begin{problem}[Wheeden \& Zygmund {\S}5, Ex.\@ 3]
Let $\left\{f_k\right\}$ be a sequence of nonnegative measurable functions
defined on $E$. If $f_k\to f$ and $f_k\leq f$ a.e.\@ on $E$, show that
$\int_E f_k\to\int_E f$.
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-HW-Current"
%%% End:
