\chapter{Qualifying Exam, January 2011}
\begin{problem}
Let
\[
G=\SL_2(\bfZ/(5))\coloneqq
\left\{\,\begin{bmatrix}a&b\\c&d\end{bmatrix}\;\middle|\;
\text{$a,b,c,d\in\bfZ/(5)$, and $ad-bc\equiv 1\pmod{5}$}\,\right\}.
\]
\begin{enumerate}[label=(\alph*)]
\item Show $|G|=120$.
\item Show
\[
N\coloneqq\left\{\,\begin{bmatrix}1&b\\0&1\end{bmatrix}\;\middle|\;b\in\bfZ/(5)\,\right\}
\]
is a Sylow $5$-subgroup of $G$.
\item Find the number of Sylow $5$-subgroups of $G$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\begin{enumerate}[label=(\alph*)]
\item Let $G$ be a group, $H$ a subgroup of $G$ with $[G:H]=2$. Suppose $K$
  is a subgroup of $G$ of odd order. Show $K\subset H$.
\item Let $G$ be a finite group and suppose there is a sequence of
  subgroups
\[
  G_0\coloneqq G\supset G_1\supset G_2\supset\cdots\supset  G_n\coloneqq H,
\]
with $[G_i:G_{i+1}]=2$ for $i\in\{1,...,n-1\}$. Suppose $H$ has odd
order. Show $H\lhd G$.
\item Suppose $|G|=2^nm$, with $m$ odd. Suppose $G$ has a normal subgroup
  $H$ of order $m$. Show there is a sequence of subgroups $G_0\coloneqq
  G\supset G_1\supset\cdots\supset G_n\coloneqq H$, with $[G_i:G_{i+1}]=2$,
  for all $i$.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $R$ be a commutative ring with identity $1\neq 0$, and let $I$ be an
ideal of $R$. Define $\rad(I)$ to be the intersection of all maximal ideals
containing $I$, with the convention $\rad(R)=R$. Let
$\sqrt{I}\coloneqq\left\{\,r\in R\;\middle|\;\text{$r^n\in R$ for some
    $n>0$}\,\right\}$.
\begin{enumerate}[label=(\alph*)]
\item Prove $\rad(I)$ is an ideal of $R$ containing $I$.
\item Prove $\sqrt{I}\subset\rad(I)$.
\item Let $F$ be a field, set $R=F[X]$, and let $I=(f)$, for some nonzero
  polynomial $f(X)\in R$. Describe $\rad(I)$ in this intstance.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $S$ be the subring of $\bfC[X]\times\bfC[Y]$ consisting of pairs
$(f,g)$ with $f(0)=g(0)$.
\begin{enumerate}[label=(\alph*)]
\item Let $\varphi\colon\bfC[X,Y]\to S$ be defined by $\varphi(h)=(f,g)$,
  where $f(X)=h(x,0)$, and $g(Y)=g(0,Y)$. Prove $\varphi$ is a surjective
  homomorphism.
\item Prove $\bfC[X,Y]/(X,Y)\cong S$.
\item Use (b) to describe the prime ideals of $S$. Be sure to justify your
  answer.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $p$ be a prime, let $F=\bfF_p$ be the field of $p$ elements and
$K=\bfF_{p^{10}}$ be the unique extension of $F$ with $p^{10}$ elements.
\begin{enumerate}[label=(\alph*)]
\item Find all subfields of $K$. Make sure to justify your answer.
\item Find a formula for the number of monic irreducible polynomials of
  degree $10$ in $F[X]$. Justify your answer.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Let $f(X)=(X^2-3)(X^3-7)\in\bfQ[X]$. Let $K$ be the splitting field of
$f(X)$ over $\bfQ$.
\begin{enumerate}[label=(\alph*)]
\item Find the degree of $K$ over $\bfQ$.
\item Classify the Galois group $\Gal(K/\bfQ)$.
\item Find all subfields $E$ of $K$ so that $E/\bfQ$ is a quadratic
  extension.
\end{enumerate}
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA553-Quals"
%%% End:
