\chapter{MA571 (Midterm 2015)}
\begin{problem}
Prove that a function to a product space is continuous if and only if its
components are.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that a subspace is closed if and only if it contains all of its limit
points.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that the projection maps for a product are open maps.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that $\partial A=\emptyset$ if and only if $A$ is open and closed.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that a metric space satisfies the 1st countability axiom.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that $\bfR^\omega$ is not metrizable in the box topology.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Show that the diagonal map is not continuous in the box topology, but it is
in the product topology.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove the sequence lemma.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Give an example of a surjective map of spaces that is not a quotient map.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that if $f_n$ is a sequence of functions $X\to\bfR$ considered as
elements of $X^{\bfR}$ with the product topology, then $f_n\to f$ if and
only if for each $x\in X$ the sequence $f_n(x)$ converges to the point
$f_n(x)$.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Prove that if $f_n$ is a sequence of functions $X\to\bfR$ considered as
elements of $X^{\bfR}$ with the topology induced by the uniform metric
$\bar\rho$, then $f_n\to f$ if and only if the sequence of functions
$f_n$ converges uniformly to the point $f$. (Recall that $f_n\colon X\to
Y$, with $Y$ a metric space, uniformly converges to $f$ if for any
$\varepsilon>0$ there exists an integer $N$ such that for all $n>N$ and
$x\in D$, $d_y(f_n(x),f(x))<\varepsilon$.)
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
Give an example of a surjective map of spaces that is not a quotient map.
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}

\begin{problem}
\end{problem}
\begin{proof}
\end{proof}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA571-Quals"
%%% End:
