\chapter{MA 544 (Spring 2016)}
\thispagestyle{empty}
\setcounter{exercise}{0}
\setcounter{equation}{0}
\section{Exam 1 Prep}
\begin{problem}
Let $E\subset\bbR^n$ be a measurable set, $r\in\bbR$ and define the set
$rE=\left\{\,r\bfx:\bfx\in E\,\right\}$. Prove that $rE$ is
measurable, and that $|rE|=|r|^n|E|$.
\end{problem}
\begin{proof}
Define a map $T\colon\bbR^n\to\bbR^n$ by $T\bfx\coloneqq r\bfx$. Note
that $T$ is
\href{https://en.wikipedia.org/wiki/Lipschitz_continuity}{\emph{Lipschitz
    continuous}} since for any $\bfx,\bfy\in\bbR^n$, the equality
\begin{equation}
\label{eq:sp16:1:1}
|T\bfx-T\bfy|=|r\bfx-r\bfy|=|r||\bfx-\bfy|
\end{equation}
is satisfied. By Theorem 3.33 from \cite[\S 3.3,
p.\@55]{book:wheeden-zygmund}, the image of $E$ under $T$ is
measurable. Moreover, by Theorem 3.35 \cite[\S 3.3, p.\@
56]{book:wheeden-zygmund}, since $T$ is linear, it follows that
$|T(E)|=|{\det T}||E|$ where $\det T=|r|^n$. Lastly, we note that the image
of $E$ under $T$ is precisely the set $rE$ so that $|T(E)|=|rE|=|r|^n|E|$,
as was to be shown.
\end{proof}

\begin{problem}
Let $\left\{ E_k \right\}$, $k\in\bbN$ be a collection of measurable
sets. Define the set
\[
\liminf_{k\to\infty} E_k
=\bigcup_{k=1}^\infty\left(\bigcap_{n=k}^\infty E_n\right).
\]
Show that
\[
\left|\liminf_{k\to\infty} E_k\right|\leq\liminf_{k\to\infty}\left|E_k\right|.
\]
\end{problem}
\begin{proof}
Following the style of \cite[\S 1.1, p.\@ 2]{book:wheeden-zygmund},
particularly, the sets defined after the introduction of equation (1.1),
set
\begin{equation}
\label{eq:sp16:1:2}
V_k\coloneqq\bigcap_{\ell=k}^\infty E_\ell.
\end{equation}
Note that the collection of sets $\{V_k\}$ forms an increasing
sequence, that is, if $\bfx\in V_k$ then, by \eqref{eq:sp16:1:2}, $\bfx$ is
in the intersection $E_k\cap\bigl(\bigcap_{\ell=k+1}E_\ell\bigr)$, but, by
\eqref{eq:sp16:1:2}, $\bigcap_{\ell=k+1}E_\ell=V_{k+1}$ thus, $\bfx$
is in $V_{k+1}$ so $V_{k+1}\supset V_k$. Hence, we have $V_k\nearrow\liminf
E_k$.

Now, consider the sequence $\{|V_k|\}$ formed by the Lebesgue measure of
the $V_k$. By Theorem 3.26 from \cite[\S 3.3, p.\@
51]{book:wheeden-zygmund}, since $V_k\nearrow\liminf E_k$,
\begin{equation}
  \label{eq:sp16:1:3}
\lim_{k\to\infty}|V_k|=
\lim_{k\to\infty}\left|\bigcap_{\ell=k}^\infty E_\ell\right|=
\left|\liminf_{k\to\infty} E_k\right|.
\end{equation}
But note that, by the monotonicity of the Lebesgue measure, we have
\begin{equation}
  \label{eq:sp16:1:4}
\left|\bigcap_{\ell=k}^\infty E_\ell\right|\leq |E_k|,
\end{equation}
so, by properties of the $\liminf$, in particular, by Theorem 19(v) from
\cite[\S 1.5, p.\@ 23]{book:royden}, we have
\begin{equation}
\label{eq:sp16:1:5}
\limsup_{k\to\infty}|V_k|\leq\liminf_{k\to\infty}|E_k|.
\end{equation}
Hence, by \eqref{eq:sp16:1:3} and Proposition 19 (iv), since the sequence
$\{|V_k|\}$ converges and is equal to the measure of $\liminf E_k$, by
\eqref{eq:sp16:1:5}, we have
\begin{equation}
\label{eq:sp16:1:6}
\left|\liminf_{k\to\infty} E_k\right|\leq\liminf_{k\to\infty}|E_k|
\end{equation}
as was to be shown.
\end{proof}

\begin{problem}
Consider the function
\[
F(x)=
\begin{cases}
|B(\mathbf{0},x)|&x>0\\
0&x=0
\end{cases}.
\]
Here
$B(\mathbf{0},r)=\left\{\,\bfy\in\bbR^n:|\bfy|<r\,\right\}$. Prove
that $F$ is monotonic increasing and continuous.
\end{problem}
\begin{proof}
% That $F$ is increasing is immediate from the monotonicity of the outer
% measure since for $x<x'$ we have $B(\mathbf{0},x)\subset B(\mathbf{0},x')$
% so, by 3.2, we have
% \[
% F(x)\left|B(\mathbf{0},x)\right|\leq\left|B(\mathbf{0},x')\right|=F(x')
% \]
% as desired.

% To see that $F$ is continuous, we will prove the following lemma
% \begin{lemma}
% \label{lem:one}
% For any $x>0$, $xB(\mathbf{0},1)=B(\mathbf{0},x)$.
% \end{lemma}
% \begin{proof}[Proof of lemma]
% \renewcommand\qedsymbol{$\clubsuit$}
% If $\bfy\in xB(\mathbf{0},1)$ then $\bfy=x\bfy'$ for $\bfy'\in
% B(\mathbf{0},1)$. Thus, $|\bfy'|=|\bfy|/x<1$ so $|\bfy|<x$ implies that
% $\bfy\in B(\mathbf{0},x)$. Hence, we have the containment
% $xB(\mathbf{0},1)\subset B(\mathbf{0},x)$.

% On the other hand, if $\bfy\in B(\mathbf{0},x)$ then $|\bfy|<x$ so
% $\left|\bfy/x\right|<1$. Hence, $\bfy/x\in B(\mathbf{0},1)$ so
% $x(\bfy/x)=\bfy\in B(\mathbf{0},x)$. Thus, $B(\mathbf{0},x)\subset
% xB(\mathbf{0},)$ and equality holds.
% \end{proof}
% In light of Lemma \ref{lem:one} and 3.35, for $x>0$, we have
% \begin{equation}
% \label{eq:scale-map}
% F(x)=\left|B(\mathbf{0},x)\right|
% =\left|xB(\mathbf{0},1)\right|
% =x^n\left|B(\mathbf{0},1)\right|.
% \end{equation}
% It is clear that $F$ is continuous on the interval $[0,\infty)$ since $F$
% is a polynomial in $x$.
Define the linear map $T\colon[0,\infty)\times\bbR^n\to\bbR^n$ by
$T(r)\bfx\coloneqq r\bfx$. We claim that
$B(\mathbf{0},r)=T(r,B(\mathbf{0},1))$. To reduce notation, set
$B_1\coloneqq B(\mathbf{0},1)$ and $B_r\coloneqq B(\mathbf{0},r)$.
\begin{proof}[Proof of claim]
\renewcommand{\qedsymbol}{$\clubsuit$}
$\subset$: Let $\bfx\in B_r$. Then $|\bfx|<r$ so $|\bfx|/r<1$. Thus,
$|\bfx|/r\in B_1$ so it is in the image of $B_1$ under the map
$T(r,\blank)$.

$\supset$: On the other hand, suppose $\bfx\in T(r,B_1)$. Then
$\bfx=r\bfy$ for some $\bfy\in B_1$. Then, since $|\bfy|<1$,
$|\bfx|=r|\bfy|<r$ so $\bfx\in B_r$.
\end{proof}

From the claim, we see that $F(x)=|T(x,B(\mathbf{0},1))|$ which, by Problem
1, is nothing more that the polynomial $|B_1|x^n$. It is clear,
from this equivalence, that $F$ is monotonically increasing: Take
$x,y\in[0,\infty)$ such that $x<y$, then $x^n<y^n$ so
\begin{equation}
\label{eq:sp16:1:7}
F(x)=|B_1|x^n<|B_1|y^n=F(y).
\end{equation}
Thus, $F$ is monotonically increasing.

In the argument above, since $F(x)=|B_1|x^n$ is a polynomial in
$[0,\infty)$ (and polynomials are continuous on $\bbR$) $F$ is continuous
on $[0,\infty)$.
\end{proof}

\begin{problem}
Let $f\colon\bbR\to\bbR$ be a function. Let $C$ be the set of all points
at which $f$ is continuous. Show that $C$ is a set of type $G_\delta$.
\end{problem}
\begin{proof}
From the topological definition of continuity, $f$ is continuous at $x\in
C$ if and only if for every neighborhood $U$ of $f(x)$, the preimage
$f^{-1}(U)$ is a neighborhood of $x$. Now,
\end{proof}
Let $x\in C$. Then, by the definition of continuity, for every natural
number $n>0$ there exists $\delta>0$ such that $|x-x'|<\delta$ implies
\begin{equation}
\label{eq:continuity-2n}
\left|f(x)-f(x')\right|<\frac{1}{2n}.
\end{equation}
Let $x'',x'\in B(x,\delta)$. Then, by the triangle inequality, we have
\begin{equation}
\label{eq:n-estimates}
\begin{aligned}
|f(x')-f(x)''|={}&\left|f(x')-f(x)-(f(x'')-f(x))\right|\\
              \leq{}&\left|f(x')-f(x)\right|+\left|f(x'')-f(x)\right|\\
              <{}&\frac{1}{2n}+\frac{1}{2n}\\
              ={}&\frac{1}{n}.
\end{aligned}
\end{equation}
In view of these estimates, define the set
\begin{equation}
\label{eq:a-n}
A_n=
\left\{\,
x\in\bbR :\text{there exists $\delta>0$ such that $x',x''\in B(x,\delta)$ implies $\left|f(x')-f(x'')\right|<\frac{1}{n}$ }
\,\right\}.
\end{equation}
Good Lord, that was a long definition! We claim that
$C=\bigcap_{n=1}^\infty A_n$ and that $A_n$ is open for all $n$.

First, let us show that $C=\bigcap_{n=1}^\infty A_n$. Let $x\in C$. Then for
every $n>0$, there exists $\delta>0$ such that $|x-x'|<\delta$ implies
$|f(x)-f(x')|<1/n$. Thus, $x\in A_n$ for all $n$ so $x\in\bigcap A_n$. On
the other hand, if $x\in\bigcap A_n$ for every $n>0$, there exists
$\delta>0$ such that $|x-x'|<\delta$ implies $|f(x)-f(x')|<1/n$. Fix
$\varepsilon>0$. By the Archimedean principle, there exists $N>0$ such that
$\varepsilon>1/N$. Then, since $x\in A_N$ it follows that for some
$\delta'>0$, $|x-x'|<\delta'$ implies
$\left|f(x)-f(x')\right|<1/N<\varepsilon$. Thus, $x\in C$ and we conclude
that $C=\bigcap_{n=1}^\infty A_n$.

Lastly, we show that $A_n$ is open. Let $x\in A_n$. Then there exists
$\delta>0$ such that $|x-x'|<\delta$ implies $|f(x)-f(x')|<1/n$. In
particular, this means that $B(x,\delta)\subset A_n$ for any $x'\in
B(x,\delta)$ satisfies $|f(x)-f(x')|<1/n$. Thus, $A_n$ is open and we
conclude that $C=\bigcap_{n=1}^\infty A_n$ is a $G_\delta$ set.
\begin{problem}
Let $f\colon\bbR\to\bbR$ be a function. Is it true that if the sets
$\left\{\,f=r\,\right\}$ are measurable for all $r\in\bbR$, then $f$ is
measurable?
\end{problem}
\begin{proof}
No. Recall that, by definition, or 4.1, $f$ is measurable if and only if
$\left\{\,f>a\,\right\}$ for all $a\in\bbR$.
\end{proof}

\begin{problem}
Let $\left\{f_k\right\}_{k=1}^\infty$ be a sequence of measurable functions
on $\bbR$. Prove that the set
$\left\{\,x:\text{$\lim_{k\to\infty} f_k(x)$ exists}\,\right\}$
is measurable.
\end{problem}
\begin{proof}
The idea here should be to rewrite
\begin{equation}
  \label{eq:measurable-lim-set}
E=\left\{\,x:\text{$\lim_{k\to\infty} f_k(x)$ exists}\,\right\}
\end{equation}
as a countable union/intersection of measurable sets. Let $x\in E$. By the
Cauchy criterion, for every $N>0$ there exists a positive integer
$M$ such that $m,n\geq M$ implies
$\left|f_n(x)-f_m(x)\right|<1/N$. With this in mind, define
\begin{equation}
  \label{eq:countable-lim-set}
E_N=
\left\{\,
x:\text{there exists $M$ such that $m,n\geq M$ implies $\left|f_n(x)-f_m(x)\right|<\frac{1}{N}$}
\,\right\}.
\end{equation}
Then, like for Problem 1.4, it is not too hard to see that the $E_n$'s are
open and that $E=\bigcap_{n=1}^\infty E_n$. Thus, $E$ is a $G_\delta$ set
and therefore measurable.
\end{proof}

\begin{problem}
A real valued function $f$ on an interval $[a,b]$ is said to be
\emph{absolutely continuous} if for every $\varepsilon>0$, there exists a
$\delta>0$ such that for every finite disjoint collection
$\left\{(a_k,b_k)\right\}_{k=1}^N$ of open intervals in $(a,b)$ satisfying
$\sum_{k=1}^Nb_k-a_k<\delta$, one has
$\sum_{k=1}^N\left|f(b_k)-f(a_k)\right|<\varepsilon$. Show that an
absolutely continuous function on $[a,b]$ is of bounded variation on
$[a,b]$.
\end{problem}
\begin{proof}
Suppose $f\colon[a,b]\to\bbR$ is absolutely continuous. Then for fixed
$\varepsilon=1$, there exists a $\delta>0$ such that for every finite
disjoint collection $\left\{ (a_kb_k) \right\}_{k=1}^N$ of open intervals
in $(a,b)$ satisfying $\sum_{k=1}^Nb_k-a_k<\delta$, we have
$\sum_{k=1}^N\left|f(b_k)-f(a_k)\right|<\varepsilon$. Let
$\Gamma=\left\{x_k\right\}_{k=1}^N$ be a partition of $[a,b]$ into closed
intervals such that $x_{k+1}-x_k<\delta$, then by absolute continuity we have
\begin{equation}
\label{eq:absolute-continuity-variation}
\begin{aligned}
V[f;\Gamma]
={}&\sum_{k=1}^N\left|f(x_{k+1})-f(x_k)\right|\\
<{}&1.
\end{aligned}
\end{equation}
Thus, $f$ is b.v.\@ on $[a,b]$.
\end{proof}

\begin{problem}
Let $f$ be a continuous function from $[a,b]$ into $\bbR$. Let
$\chi_{\{c\}}$ be the characteristic function of a singleton
$\left\{c\right\}$, that is, $\chi_{\{c\}}(x)=0$ if $x\neq c$ and
$\chi_{\{c\}}(c)=1$. Show that
\[
\int_a^b f d \chi_{\{c\}}=
\begin{cases}
0&\text{if $c\in(a,b)$}\\
-f(a)&\text{if $c=a$}\\
f(a)&\text{if $c=b$}
\end{cases}.
\]
\end{problem}
\begin{proof}
\end{proof}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA544-Quals"
%%% End:
