\subsection{McClure: Winter 2012}
\setcounter{exercise}{0}

\begin{problem}
  Let \(X\) be a topological space. Recall that a subset of \(X\) is
  \emph{dense} if its closure is \(X\). Prove that the intersection of two
  dense open sets is dense.
\end{problem}
\begin{solution}
  Suppose \(U\) and \(V\) are open dense subsets of \(X\). We will show
  that \(U\cap V\) is dense in \(X\), i.e., \(\overline{U\cap V}=X\). To
  that end, we will show that for any point \(x\in X\), for any
  neighborhood \(W\) of \(x\), \(W\cap(U\cap V)\neq\emptyset\). Therefore,
  let \(x\in X\). Let \(W\) be a neighborhood of \(x\). Then, since \(U\)
  is dense in \(X\), \(W\cap U\neq\emptyset\). Let \(y\in W\cap U\). Then,
  since \(U\) and \(V\) are open, \(U\cap V\) is open so \(U\cap V\) is a
  neighborhood of \(y\). Moreover, since \(V\) is dense in \(X\),
  \((W\cap U)\cap V\neq\emptyset\). Now, since intersection is associative,
  \((W\cap U)\cap V=W\cap(U\cap V)\neq\emptyset\). Thus,
  \(x\in\overline{U\cap V}\) and we have \(\overline{U\cap V}=X\) as
  desired.
\end{solution}

\begin{problem}
  Let \(X\) be a set with two elements \(\{a,b\}\). Give \(X\) the
  \emph{indiscrete} topology. Give \(X\times\bbR\) the product
  topology. Let \(A\subset X\times\bbR\) be
  \((\{a\}\times[0,1])\cup(\{b\}\times(0,1))\). Prove that \(A\) is
  compact.

  You may use the fact that a set is compact if every covering by
  \emph{basic} open sets has a finite subcovering.
\end{problem}
\begin{solution}
  Let \(\calU\) be an open cover of \(A\) by basic open sets. Then each
  \(U\in\calU\) is of the form \(\{a,b\}\times V\) where \(V\) is an open
  subset of \(\bbR\). Then, the \(V\)'s, i.e., \(\pi_2(U)\) where
  \(\pi_2\colon X\times\bbR\to\bbR\) is an open map by previous work, form
  open cover of \([0,1]\) (since \(\bigcup_{U\in\calU} U\supset A\), we
  must have \(\bigcup_{U\in\calU}\pi_2(U)\supset[0,1]\). Now, since
  \([0,1]\) is compact in \(X\) there is a finite collection of the
  \(V\)'s, say \(\left\{V_1,\dotsc,V_n\right\}\), that cover
  \([0,1]\). Call \(U_i\) the element of \(\calU\) such that
  \(\pi_2(U_i)=V_i\). Then the \(U_i\)'s form a finite subcover of
  \(A\). Thus, \(A\) is compact.
\end{solution}

\begin{problem}
  Let \(B^2\) be the disk
  \[
    \left\{\,(x,y)\in\bbR^2:x^2+y^2\leq 1\,\right\}.
  \]
  Let \(S^1\) be the circle
  \[
    \left\{\,(x,y)\in\bbR^2:x^2+y^2=1\,\right\}.
  \]
  Prove that there is an equivalence relation \(\sim\) such that \(B^2\) is
  homeomorphic to \((S^1\times[0,1])/{\sim}\). As port of your solution
  explain how you are using one or more properties of the quotient
  topology.
\end{problem}
\begin{solution}
  Such an equivalence relation is called the cone of \(S^1\). We define it
  as follows, let \((x,y,z),(x',y',z')\in S^1\times[0,1]\) then we say
  \((x,y,z)\sim(x',y',z')\) if and only if \((x,y)=(x',y')\) or
  \(z=z'=0\). We shall take it on faith that \(\sim\) is in fact an
  equivalence relation (we may return to this if time permits).

  By the UMP of the quotient space, we need to find a continuous surjection
  \(f\colon S^1\times[0,1]\to B^2\) that preserves the equivalence relation
  \(\sim\). So consider the map \(f(x,y,z)=(zx,zy)\). This map is
  continuous by Theorem 18.4 since \(\pi_1\circ f(x,y,z)=zx\) is
  multiplication on \(\bbR\) and similarly for \(\pi_2\circ
  f(x,y,z)\). Moreover, this map preserves the equivalence relation: let
  \((x,y,z)\sim(x',y',z')\) then \((x,y,z)=(x',y',z')\) in which case
  \[
    f(x,y,z)=(zx,zy)=(z'x',z'y')=f(x',y',z')
  \]
  or \(z=z'=0\) so
  \[
    f(x,y,0)=(0\cdot x,0\cdot y)=(0,0)=(0\cdot x',0\cdot y')=f(x',y',0).
  \]
  In either case, we have \(f(x,y,z)=f(x',y',z')\). Thus, by the UMP of the
  quotient space, the induced map
  \(f'\colon (S^1\times[0,1])/{\sim}\to B^2\) is continuous.

  Now, since \(S^1\times[0,1]\) is closed and bounded, by Heine--Borel,
  \(S^1\times[0,1]\) is a compact subset of \(\bbR^3\). Therefore,
  \((S^1\times[0,1])/{\sim}\) is compact. Since \(B^2\subset\bbR^2\) is
  Hausdorff, it suffices to show, by Theorem 26.6, that \(f\) is bijective.

  It is eassy to see that \(f\) is surjective since for any point
  \((x,y)\neq(0,0)\) in \(B^2\), \(\sqrt{x^2+y^2}\leq 1\) so letting
  \(z=\sqrt{x^2+y^2}\), \(x'=x/\sqrt{x^2+y^2}\), and
  \(y'=y/\sqrt{x^2+y^2}\) we have
  \[
    f(x',y',z)=\sqrt{x^2+y^2} \left(\frac{x}{\sqrt{x^2+y^2}},
      \frac{y}{\sqrt{x^2+y^2}}\right)=(x,y).
  \]
  And, trivially, if \((x,y)=0\), we have \(\varphi(x,y,0)=0\) for any
  \((x,y)\in S^1\).

  To see that it is injective, merely note that, by the definition of
  \(f\), \(f(x,y,z)=f(x',y',z')\) if and only if \((x,y,z)=(x',y',z')\) or
  \(z=z'=0\) which precisely means that \((x,y,z)\sim(x',y',z')\). Thus,
  \(f\) is injective.

  It follows that \((S^1\times[0,1])/{\sim}\approx B^2\).
\end{solution}

\begin{problem}
  Let \(X\) be a set with \(2\) elements \(\{a,b\}\). Give \(X\) the
  \emph{discrete} topology. Let \(Y\) be any topological space. Recall that
  \(\scrC(X,Y)\) denotes the set of continuous functions from \(X\) to
  \(Y\), with the compact-open topology. Prove that \(\scrC(X,Y)\) is
  homeomorphic to \(Y\times Y\) (with the product topology).
\end{problem}
\begin{solution}
  Consider the map \(F\colon\scrC(X,Y)\to Y\times Y\) given by
  \(F(f)=(f(a),f(b))\). This map is continuous by Theorem 18.4, since
  \(\pi_1(F)\) and \(\pi_2(F)\) are, respectively, the evaluation of \(f\)
  at \(a\) and the evaluation of \(f\) and \(b\), both of which are
  continuous because under the compact-open topology. This map is clearly
  surjective since for any \((y_1,y_2)\in Y\times Y\) we may define the
  function \(f(a)= y_1\) and \(f(b)= y_2\) which is continuous sinced \(X\)
  has the discrete topology. Moreover, \(F\) is injective since if
  \((f(a),f(b))=(g(a),g(b))\) then \(f(x)=g(x)\) for all \(x\in X\) hence,
  \(f=g\). Therefore, to show that \(F\) is a homeomorphism, it suffices to
  show that \(F\) is an open map.

  Now it suffices to find a continuous inverse. For any
  \((y_1,y_2)\in Y\times Y\), define the map
  \(g\colon Y\times Y\to\scrC(X,Y)\).
  \[
    g(y_1,y_2)= f (y)=\begin{cases}
      a&\text{if \(y=y_1\)}\\
      b&\text{if \(y=y_2\)}.
    \end{cases}
  \]

\end{solution}

\begin{problem}
  Let \(X\) and \(Y\) be homotopy-equivalent topological spaces. Suppose
  that \(X\) is path-connected. Prove that \(Y\) is path-connected.
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}
  Suppose that \(X\) is a wedge of two circles: that is, \(X\) is a
  Hausdorff space which is a union of two subspaces \(A_1\) and \(A_2\)
  such that \(A_1\) and \(A_2\) are each homeomorphic to \(S^1\) and
  \(A_1\cap A_2\) is a single point \(p\).

  Use the Seifert--van Kampen theorem to calculate \(\pi_1(X,p)\). You
  should state what deformation retractions you are using, but you do not
  have to give formulas for them.
\end{problem}
\begin{solution}
\end{solution}

\begin{problem}
  Let \(p\colon E\to B\) be a covering map. Let \(A\) be a connected space
  and let \(a\in A\). Prove that if two continuous functions
  \(\alpha,\beta\colon A\to E\) have a property that \(\alpha(a)=\beta(a)\)
  and \(p\circ\alpha=p\circ\beta\) then \(\alpha=\beta\).

  For partial credit, you may assume that \(p\) is the standard covering
  map from \(\bbR\) to \(S^1\).
\end{problem}
\begin{solution}
\end{solution}

Here's an extra problem I felt like doing since I thought it might be on
the exam:
\begin{problem*}
  \begin{theorem*}[Munkres, Theorem 18.4]
    Let \(f\colon A\to X\times Y\) be given by the equation
    \(f(a)=(f_1(a),f_2(a))\). Then \(f\) is continuous if and only if
    \(f_1\colon A\to X\) and \(f_2\colon A\to Y\) are continuous.
  \end{theorem*}
\end{problem*}
\begin{solution}
  Let \(\pi_1\colon X\times Y\to X\) and \(\pi_2\colon X\times Y\to Y\) be
  projections onto the 1st and 2nd factors, respectively. These maps are
  continuous and open by previous work. Now, for every \(a\in A\) we have
  \[
    \pi_1(f(a))=f_1(a)\qquad\text{and}\qquad \pi_2(f(a))=f_2(a).
  \]
  Therefore, if \(f\) is continuous, then \(f_1\) and \(f_2\) are the
  composites of the continuous functions above therefore, are continuous.

  Conversely, suppose that \(f_1\) and \(f_2\) are continuous. By Lemma C,
  it suffices to show that for each basic open set
  \(U\times V\subset X\times Y\), the preimage \(f^{-1}(U\times V)\) is
  open. But \(a\in f^{-1}(U\times V)\) if and only if
  \(f(a)\in U\times V\), if and only if \(f_1(a)\in U\) and
  \(f_2(a)\in V\). Thus, \(f^{-1}(U\times V)=f^{-1}(U)\cap f^{-1}(V)\)
  which is open in \(A\) since \(U\) is open in \(X\) and \(V\) is open in
  \(Y\) and \(f_1,f_2\) are continuous.
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA571-Quals"
%%% End:
