\section{MA 553 Spring 2016}
% \thispagestyle{empty}
This is material from the course MA 533 as it was taught in the spring of
2016.%
\bigskip
\subsection{Homework}
Most of the homework is Ulrich original (or as original as elementary
exercises in abstract algebra can be). However, an excellent resource and
one that I will often quote on these solutions is \cite{hungerford}. Other
resources include \cite{dummit-foote} and (to a lesser extent)
\cite{herstein}. I may also cite Milne's \emph{Group Theory}, \emph{Field
  Theory}, and \emph{Commutative Algebra: A Primer} notes, respectively,
\cite{milneGT}, \cite{milneFT}, and (no reference for the last). Unless
otherwise stated, whenever we quote a result, e.g., Theorem 1.1, it is
understood to come from Hungerford's \emph{Algebra}.

Throughout these notes

\begin{tabular}{cl}
  $\bbR$   & is the set of real numbers\\
  $\bbC$   & is the set of complex numbers\\
  $\bbQ$   & is the set of rational numbers\\
  $\bbF_q$ & is the finite field of order $q=p^n$ for some prime $p$\\
  $\bbZ$   & is the set of the integers\\
  $\bbN$   & is the set of the natural numbers $1,2,\dotsc$\\
  $k$   & is used to denote the base field with
             characteristic $\Char k$\\
  $K,E,L$& is used to denote field extensions over the base field
                    $k$\\
  $Z_n$    & is the cyclic group of order $n$ not necessarily equal
             (but isomorphic) to $\bbZ/p\bbZ$\\
  $S_n$    & is the symmetric group on $\{1,\dotsc,n\}$\\
  $A_n$    & is the alternating group on $\{1,\dotsc,n\}$\\
  $D_n$    & is the dihedral group of order $n$\\
  $A\setminus B$ & is the set difference of $A$ and $B$, that is, the
                        complement of $A\cap B$ in $A$\\
  $X\cong Y$ & means $X$ and $Y$ are isomorphic as groups, rings,
                $R$-modules, or fields
\end{tabular}

\newpage

\subsubsection{Homework 1}
\setcounter{exercise}{0}
\setcounter{equation}{0}

\begin{problem}
  Let $G$ be a group, $a\in G$ an element of finite order $m$, and $n$ a
  positive integer. Prove that
  \[
    |a^n|=\frac{m}{(m,n)}.
  \]
\end{problem}
\begin{solution}
  Let $\ell$ denote the order of $a^n$. Then $\ell$ is the minimal power of
  $a^n$ such that ${(a^n)}^\ell=e$. Now, observe that
    \begin{align*}
      {(a^n)}^{m/(m,n)}
      &=a^{nm/(m,n)}\\
      &=a^{mn/(m,n)}\\
      &={(a^m)}^{n/(m,n)}\\
      &=e^{n/(m,n)}\\
      &=e.
    \end{align*}
  Thus $\ell\leq m/(m,n)$.

  On the other hand, by Theorem 3.4 (iv) since ${(a^n)}^\ell=a^{n\ell}=e$
  and the order of $a$ is $m$, $m\mid n\ell$ or, equivalently, $mk=n\ell$
  for some $k\in\bbZ^+$. Now, since $(m,n)\mid m$ and $(m,n)\mid n$, we can
  represent $m$ and $n$ as the products $(m,n)m'$ and $(m,n)n'$,
  respectively. Now, note that $m'=m/(n,m)$ so we must show that
  $m'\leq\ell$. Putting all of this together, we have $mk$
  \[
    mk=(m,n)m'k=(m,n)n'\ell=n\ell
  \]
  so
  \[
    m'k=n'\ell.
  \]
  Thus $m'\mid n'\ell$ so either $m'\mid n'$ or $m'\mid\ell$. But since we
  factored the $(m,n)$ from $m$ and $n$, it follows that $(m',n')=1$ so
  $m'\mid \ell$. Therefore $m'\leq\ell$ and equality holds, that is,
  $\ell=m/(m,n)$.
\end{solution}

\begin{problem}
  Let $G$ be a group, and let $a$, $b$ be elements of finite order $m$, $n$
  respectively. Show that if $ba=ab$ and
  $\langle a\rangle\cap\langle b\rangle=\{e\}$, then $|ab|=mn/(m,n)$.
\end{problem}
\begin{solution}
  Let $\ell$ denote the order of $ab$. Now, playing around with powers of
  $ab$, we have
  \begin{align*}
    (ab)^{n}
    &=a^nb^n\\
    &=a^n\\
    &\neq e
  \end{align*}
  since the order of $a$ is $m$ and $n<m$. Thus, by Problem 1,
  $|a^n|=m/(m,n)$ so $|ab|=mn/(m,n)$.
\end{solution}

\begin{problem}
  Let $G$ be a group and $H$, $K$ normal subgroups with $H\cap
  K=\{e\}$. Show that
  \begin{enumerate}[label=(\alph*),noitemsep]
  \item $hk=kh$ for every $h\in H$, $k\in K$.
  \item $HK$ is a subgroup of $G$ with $HK\cong H\times K$.
  \end{enumerate}
\end{problem}
\begin{solution}
  (a) Suppose that $H$ and $K$ are normal in $G$. Then, for every $g\in G$,
  $gh=hg$ and $gk=kg$ for any $h\in H$, $k\in K$. In particular, since
  $H\subseteq G$, $h\in G$ so $hk=kh$.
  \\\\
  (b) Consider the subset $HK$ of $G$ consisting of all products $hk$ where
  $h\in H$, $k\in K$. First, we show that $HK$ is closed under
  multiplication: Pick $h_1k_1,h_2k_2\in HK$ then
  $h_1k_1h_2k_2=h_1(k_1k_2)h_2=h_1h_2(k_1k_2)$ is in $HK$ since $h_1h_2\in
  H$, $k_1k_2\in K$. Moreover, since $e\in H$ and $e\in K$, $ee=e\in
  HK$. Lastly, given $hk\in HK$,
  $hkh^{-1}k^{-1}=(hkh^{-1})k^{-1}=kk^{-1}=e$ so $HK$ is closed under
  taking inverses. Thus, $HK$ is a subgroup of $G$.

  To see that $HK\cong H\times K$, consider the map
  $\varphi\colon HK\to (HK/K)\times(HK/H)$ given by
  $\varphi(hk)=(\pi_K(h),\pi_H(k))$ where $\pi_H\colon HK\to HK/H$
  and $\pi_K\colon HK\to HK/K$ are quotient maps. By the first (or second)
  isomorphism theorem, $H\cong HK/H$ and $K\cong HK/H$ so $HK\cong
  H\times K$.
\end{solution}

\begin{problem}
  Show that $A_4$ has no subgroup of order $6$ (although $6\mid 12=|A_4|$).
\end{problem}
\begin{solution}
  We proceed by contradiction. Suppose that $A_4$ has a subgroup of order
  $6$, call it $H$. Then, we claim that $H$ must contain all elements
  $\sigma^2$ where $\sigma\in A$.
  \begin{proof}[Proof of claim]
    Since $|H|=6$, $[A_4:H]=2$ which implies that $H$ is must be a normal
    subgroup of $A_4$. Now, consider the collection of $G/H$ of
    right-cosets of $H$ in $G$. By Theorem 5.4, $G/H$ is a group with order
    $|G/H|=2$ so either $\bar\sigma=\bar e$ or ${\bar\sigma}^2=\bar
    e$. Thus, $\sigma^2\in H$.
  \end{proof}
  Thus, $H$ must contain all of the squares in $A_4$. However, counting all
  of the elements in $A_4$ and squaring them
  \begin{align*}
    (1)^2&=(1)
    &(1\,2\,3)^2&=(1\,3\,2)
    \\
    (1\,3\,2)^2&=(1\,2\,3)
    &(1\,2\,4)^2&=(1\,4\,2)\\
    (1\,4\,2)^2&=(1\,2\,4)
    &(1\,3\,4)^2&=(1\,4\,3)\\
    (1\,4\,3)^2&=(1\,3\,4)
    &(2\,3\,4)^2&=(2\,3\,4)\\
    (2\,4\,3)^2&=(2\,4\,3)
    &((1\,2)(3\,4))^2
             &=(1)\\
    ((1\,3)(2\,4))^2&=(1)&
    ((1\,4)(2\,3))^2&=(1)
  \end{align*}
  we see that there are a total of $9$ squares ($8$ nontrivial ones) which
  exceeds the order of $H$. This is a contradiction therefore, $G$ has no
  subgroup of order $6$.
\end{solution}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../MA553-Quals"
%%% End:
