\section{Ulrich}
\subsection{Ulrich: Winter 2002}
\setcounter{exercise}{0}
\setcounter{equation}{0}

\begin{problem}
  Let \(G\) be a group and \(H\) a subgroup of finite index. Show that
  there exists a normal subgroup \(N\) of \(G\) of finite index with
  \(N\subseteq H\).
\end{problem}
\begin{solution}
  Suppose \(N<G\) with \(n=[G:N]<\infty\). Let \(G\) act on \(H\) by
  translation. This action gives a homomorphism \(\varphi\colon G\to
  S_n\). Then, by the first isomorphism theorem
  \({[G:{\Ker\varphi}]}\mid{\card S_n=n!}\). Thus, \(\Ker\varphi\) is a
  normal subgroup of \(G\) with finite index.
\end{solution}

\begin{problem}
  Show that every group of order \(992\) (\(=32\cdot 31)\) is solvable.
\end{problem}
\begin{solution}
  Suppose \(\card G=992=32\cdot 31=2^5\cdot 31\). By Sylow's theorem, \(G\)
  has \(1\) or \(32\) \(31\)-Sylow subgroups. In the former case, this
  implies that there is a unique \(31\)-Sylow subgroup \(P\) and therefore
  \(P\trileq G\). Moreover, since \(\card(G/P)=2^3\), \(G/P\) is solvable
  since it is a \(p\)-group. Thus, both \(G/P\) and \(P\) are solvable
  (the latter since it is Abelian), which implies that \(G\) is solvable.

  On the other hand, if \(G\) contains \(32\) \(31\)-Sylow subgroups, then
  there are exactly \(32\cdot 31-32\cdot 30=32\) elements not of order
  \(31\). This implies that there is exactly one \(2\)-Sylow subgroup \(Q\)
  in \(G\). Again, since \(\card G/Q=31\), \(G/Q\) is solvable and \(Q\) is
  solvable since it is a \(p\)-group. Thus, \(G\) is solvable.

  In every case, \(G\) we see that is solvable.
\end{solution}

\begin{problem}
  Let \(G\) be a group of order \(56\) with a normal \(2\)-Sylow subgroup
  \(Q\), and let \(P\) be a \(7\)-Sylow subgroup of \(G\). Show that either
  \(G\simeq P\times Q\) or \(Q\simeq\bbZ/(2)\times\bbZ/(2)\times\bbZ/(2)\).
  [\emph{Hint}: \(P\) acts on \(Q\setminus\{e\}\) via conjugation. Show
  that this action is either trivial or transitive.]
\end{problem}
\begin{solution}
  Suppose \(G\) is a group of order \(56=2^3\cdot 7\) with a normal
  \(2\)-Sylow subgroup \(Q\) and let \(P\in\Syl_7(G)\). Taking the hint,
  let \(P\) act on \(Q\) by conjugation. This action gives a homomorphism
  \(\varphi\colon P\to \Aut Q\). The kernel of this action is exactly the
  centralizer \(Q\) in \(P\), \(\Ker\varphi=\rmC_P(Q)\). Considering the
  Cardinality of \(P\), either \(\Ker\varphi=P\) or
  \(\Ker\varphi=\{e\}\). In the former case, this implies that \(pq=qp\)
  for every \(p\in P\), \(q\in Q\). In particular, \(Q\) is in the
  normalizer of \(P\) and since \(\card P\mid\rmN_G(P)\), we must have
  \(\rmN_G(P)=G\). Thus, since \(P,Q\trileq G\), \(P\cap Q=\{e\}\) and
  \(PQ=G\), we have \(G\simeq P\times Q\).

  On the other hand, if \(\Ker\varphi=\{e\}\) then \(P\) acts transitively
  on \(Q\). Since conjugation is an order preserving action, and \(Q\)
  contains at least one element of order \(2\) (by Cauchy's theorem), every
  element \(q\in Q\) is of order \(2\). Now, if \(a,b\in Q\) are distinct
  nontrivial elements, \(a=a^{-1},b=b^{-1}\) and \((ab)^2=e\) implies
  \(ab=b^{-1}a^{-1}=ba\). Thus, \(Q\) must be Abelian. It follows that
  \(Q\simeq Z_2\times Z_2\times Z_2\) (since this is the only group of
  order \(8\), up to isomorphism, such that every nontrivial element has
  order \(2\)).
\end{solution}

\begin{problem}
  Let \(R\) be a commutative ring and \(\Rad(R)\) the intersection of all
  maximal ideals of \(R\).
  \begin{enumerate}[label=(\alph*)]
  \item Let \(a\in R\). Show that \(a\in\Rad(R)\) if and only if \(1+ab\)
    is a unit for every \(b\in R\).
  \item Let \(R\) be a domain and \(R[X]\) the polynomial ring over
    \(R\). Deduce that \(\Rad(R[X])=0\).
  \end{enumerate}
\end{problem}
\begin{solution}
  For part (a), \(\implies\) seeking a contradiction, suppose that \(1+ab\)
  is not a unit. By Krull's theorem, there exists a maximal ideal
  \(\frakm\) containing \(1+ab\). However, since \(a\in\frakm\) for every
  maximal ideal \(\frakm\) in \(R\), \(a\in\frakm\). This implies that
  \(ab\in\frakm\) so \(1+ab-ab=1\in\frakm\). This contradicts the
  assumption that \(\frakm\) is a maximal ideal. Thus, \(1+ab\) must have
  been a unit.

  \(\impliedby\) On the other hand, suppose \(1+ab\) is a unit for every
  \(b\in R\). If \(a\notin\Rad R\), then \(a\notin\frakm\) for some maximal
  ideal \(\frakm\). By maximality, \((a)+\frakm=R\), i.e., there exists
  \(x\in R\) and \(m\in\frakm\) such that \(ax+m=1\). Thus,
  \(m=1-ax=1+a(-x)\), but \(1+ab\) is a unit for every \(b\in R\). This
  contradicts the fact that \(\frakm\) is a maximal ideal.

  For part (b), by part (a), \(f\in\Rad(R[X])\) if and only if
  \(1+fg\) is a unit for every \(g\in R[X]\). Since the only units in
  \(R[X]\) are in \(R\), this implies that \(1+fg\in R\) for every \(g\in
  R[X]\). This is true if and only if \(f=0\) for otherwise \(1+fX\) is a
  polynomial contained in \(R\), but \(R\cap(x)=\{0\}\). Thus, \(\Rad
  R=\{0\}\).
\end{solution}

\begin{problem}
  Let \(R\) be a unique factorization domain and \(\frakp\) a prime ideal
  of \(R[X]\) with \(\frakp\cap R=0\).
  \begin{enumerate}[label=(\alph*)]
  \item Let \(n\) be the smallest possible degree of a nonzero polynomial
    in \(\frakp\). Show that \(\frakp\) contains a primitive polynomial
    \(f\) of degree \(n\).
  \item Show that \(\frakp\) is the principal ideal generated by \(f\).
  \end{enumerate}
\end{problem}
\begin{solution}
  For part (a), pick \(f\in\frakp\) of degree \(n\). Then
  \[
    f(X)=a_nX^n+\dotsb+a_1X+a_0.
  \]
  Since \(R\) is a u.f.d.\@, we can take the
  \(a=\gcd\{a_n,\dotsc,a_1,a_0\}\) and \(b_i\in R\) such that
  \(a_i=ab_i\). Then,
  \[
    g(X)=b_nX^n+\dotsb+b_1X+b_0
  \]
  is a primitive polynomial since, by construction
  \(\gcd\{b_n,\dotsc,b_1,b_0\}=1\).

  For part (b), it is clear that \((g)\subseteq\frakp\). Let \(f\in\frakp\)
  and \(F=\Frac R\). To see the reverse containment note that \(g\) is
  irreducible in \(R\) since, by Gauß's lemma, if \(g=pq\) for some
  nontrivial (\(\deg>0\)) polynomials \(p,q\in F[X]\), then there exists
  \(a,b\in F[x]\) such that \(p'=ap,q'=bq\in R[X]\) and \(g=p'q'\). By
  primality of \(\frakp\), this implies either \(p'\in\frakp\) or
  \(q'\in\frakp\). But this is an impossibility since \(g\) is of minimal
  degree in \(\frakp\). Now, let \(f\in\frakp\). Then, embedding \(R\) in
  its field of fractions \(F\), by the Euclidean algorithm there exists
  \(p,r\in F[X]\) with \(\deg r<\deg g\) or \(r=0\) such that
  \(f=pg+r\). Clearing denominators if necessary, \(r\) (or some multiple
  of it) is in \(\frakp\). Thus, \(r=0\) for otherwise, we contradict the
  minimality of \(\deg g\) in \(\frakp\). Thus, \((g)=\frakp\) as was to be
  shown.
\end{solution}

\begin{problem}
  Let \(k\) be a field of characteristic zero. Assume that every polynomial
  in \(k[X]\) of odd degree and every polynomial in \(k[X]\) of degree two
  has a root in \(k\). Show that \(k\) is algebraically closed.
\end{problem}
\begin{solution}
  We show that every polynomial \(f\in k[X]\) has a root in \(k\). Let
  \(K\) be the splitting field of \(f\). Then \([K:k]=2^\alpha m\) for some
  odd positive integer \(m\). Let \(G=\Gal(K/k)\) and let \(P\) be a
  \(2\)-Sylow subgroup of \(G\). By the Galois correspondence, \(K^P\) is a
  subfield of \(K\) of index \([G:P]=m\) over \(k\). Since every extension
  over a field of characteristic \(0\) is separable, by the primitive
  element theorem, \(K^P=K(\alpha)\) where \(\alpha\) is the root of some
  irreducible polynomial \(f\) of degree \(m\) (namely, its minimal
  polynomial). But by assumption, \(f\) has a root in \(\alpha\). Thus,
  \(k(\alpha)\subseteq k\). Thus, the degree of the splitting field must be
  \([K:k]=2^\alpha\). Thus, \(\card G=2^\alpha\) so \(G\) is has a normal
  subgroup of order \(p^k\) for every \(0\leq k\leq\alpha\). Take \(N\) of
  index \([G:N]=2\). Then, by the primitive element theorem
  \(K^N=k(\beta)\) for \(\beta\) the root of polynomial \(g\) of degree
  \(2\). Thus, \(k(\beta)\subseteq k\). Repeat this method recursively
  until \(\alpha=0\). Thus, \(k\) is algebraically closed.
\end{solution}

\begin{problem}
  Let \(k\subseteq K\) be a finite Galois extension with Galois group
  \(\Gal(K/k)\), let \(L\) be a field with \(k\subseteq L\subseteq K\), and
  set \(H=\bigl\{\,\sigma\in\Gal(K/k):\sigma(L)=L\,\bigr\}\).
  \begin{enumerate}[label=(\alph*)]
  \item Show that \(H\) is the normalizer of \(\Gal(K/L)\) in
    \(\Gal(K/k)\).
  \item Describe the group \(H/{\Gal(K/L)}\) as an automorphism group.
  \end{enumerate}
\end{problem}
\begin{solution}
  For part (a), let \(N\) denote the normalizer of \(\Gal(K/L)\) in
  \(\Gal(K/k)\). Then for any \(\sigma\in H\), \(\tau\in\Gal(K/L)\) and
  \(x\in L\) we have
  \begin{align*}
    \sigma^{-1}\circ\tau\sigma(x)
    &=\sigma^{-1}(\tau(\sigma(x)))\\
    &=\sigma^{-1}(\sigma(x))\\
    &=x.
  \end{align*}
  Thus, \(\sigma\circ\tau\sigma^{-1}\) fixes \(L\) so
  \(\sigma\circ\tau\sigma^{-1}\in\Gal(K/L)\) so \(H\subseteq N\). On the
  other hand, if \(\sigma\in N\) then we claim \(\sigma(L)=L\). Otherwise
  there exists \(x\in L\) such that \(\sigma(x)\notin L\). Since \(K\) is
  Galois over \(k\), it is Galois over \(L\). Thus, there is an element
  \(\tau\in\Gal(K/L)\) such that \(\tau\circ\sigma(x)\neq\sigma(x)\). Thus,
  \(\sigma^{-1}\circ\tau\sigma(x)\neq x\) so
  \(\sigma^{-1}\circ\tau\sigma\notin\Gal(K/L)\). This is a
  contradiction. Thus, we conclude that \(H=N\).

  For part (b), we say that \(H/{\Gal(K/L)}\) is precisely the
  automorphisms on \(L\) which do not leave \(L\).
\end{solution}

%%% Local Variables:
%%% mode: latex

%%% End:
